---
title: "Summarising the analysis of Mediterranean data using MLEbins"
author: "Andrew Edwards and Juliana Zabala"
output: pdf_document
fontsize: 12pt
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  cache = FALSE,
  # cache_path = "med-summary-4-cache/",
  # fig.path = "med-summary-4-figs-cache/",
  fig.width = 8,
  fig.height = 10
)
  # comment = "#>",
```

```{r, build, echo = FALSE, eval = FALSE}
# To build either run this line or click knit button in RStudio:
rmarkdown::render("med-summary-4.Rmd")
```

```{r, packages, echo = FALSE, cache = FALSE}
library(dplyr)
load_all()
#library(sizeSpectra2)    # Need latest version
```

Copying from `mediterranean-analysis-2/zabala-summary.Rmd`, which I hadn't
updated for `mediterranean-analysis-3/`. Committing before changing anything
further.

Need to have already run `med-analysis-4.Rmd` which takes a while to run
and saves four `*_list.rds`
files in the same directory. Or just have the three files. The files contain the
results after some exploration and iterative calculations.

## Results

TODO had originally used `aggregate_mlebins()` to make tailored list, but now
just having all results then doing such wrangling here as will keep plotting code
more general (and avoids unpacking what I packed up). Can delete this comment
when done.

```{r, loadresults}
baseline_agg_list <- readRDS("baseline_agg_list.rds")
fg_agg_list <- readRDS("fg_agg_list.rds")
ntr_agg_list <- readRDS("ntr_agg_list.rds")
res_table_agg <- readRDS("res_table_agg.rds")
```

Each is a list for that strata, with 6 objects of results of class
`determine_xmin_and_fit_mlebins`. First of each is for fitting the full
community in one go, we might want to show just the curve of that fit; the
actual data will be created anyway by the amalagamating plotting.

`res_table_agg` is the summary table that has been built up during analysis (and
users should do that by using our .Rmd as a template). Was going to rebuild it
but just what I created before for speed.

Do not really need the histogram information in each list here, and may not even
have it. So to generalise the code save strip out the histogram information if
it is exists: (could functionalise this, but a bit more transparent like this)
```{r, striphist}
if("determine_xmin_and_fit_mlebins" %in% class(baseline_agg_list[[i]])){
  baseline_agg <-   HERE strip out the histogram stuff from each, just do a
  loop. May then make into a function.

```




TODO so making `aggregated_data` only in the plotting.
data in objects).

\clearpage

```{r, aggregatetable}
knitr::kable(res_table_agg,
             digits = 2)
```

First, need a global range for the body sizes, note those include full community
that we might not look at, so may always want to come back and tweak
```{r, globalx}
x_min_global <- min(res_table_agg$x_min)
x_max_global <- max(res_table_agg$x_max)
groups <- unique(res_table_agg$Group)    # Assume we're keeping everything in
                                         # the same order, no reason not to
groups <- groups[groups != "Full community"]  # Since don't want plots for full community
```

TODO Need to summarise those results in a plot of $b$ with confidence
intervals.

## Figures -- one figure with 12 panels

Could also plot each one in turn with linear y-axis also will probably want in
an Appendix. Do panel plot first.

```{r, panelplot, fig.cap = "Summary plot of all groups (rows) and strata(columns), with a consistent x-axis range throughout, and a consistent y-axis range for each group. Data are shown as rectangles representing the range of possible values within each body-mass bin, but due to the small bins these mostly show up as lines."}
par(mfrow = c(5, 3),    # group in each row, strata in each column
    oma = c(0, 3, 3, 0),
    mar = c(3, 3, 1, 0.2))

seg_col = "black"      # bins are so small the default green does not really
                       # help

for(i in 1:length(groups)){
  # Based on calculation in plot.size_spectrum_mlebin():
  y_min_group <- 0.75 * min(c(baseline_agg[[i]]$data$count_gte_bin_min,
                              fg_agg[[i]]$data$count_gte_bin_min,
                              ntr_agg[[i]]$data$count_gte_bin_min))

  y_max_group <- max(c(baseline_agg[[i]]$data$high_count,
                       fg_agg[[i]]$data$high_count,
                       ntr_agg[[i]]$data$high_count))

  y_lim_group = c(y_min_group,
                  y_max_group)

  plot(baseline_agg[[i]],
       xlim = c(x_min_global,
                x_max_global),
       ylim = y_lim_group,
       style = "log_y_axis",
       legend_text_second_row_multiplier = 4,
       seg_col = seg_col,
       show_fit_on_top = FALSE)

  mtext(groups[i],
        side = 2,
        line = 3)

  if(i == 1){
    mtext("Baseline",
          side = 3,
          line = 1)
  }

  plot(fg_agg[[i]],
       xlim = c(x_min_global,
                x_max_global),
       ylim = y_lim_group,
       style = "log_y_axis",
       legend_text_second_row_multiplier = 4,
       seg_col = seg_col,
       ylab = "",
       show_fit_on_top = FALSE)

  if(i == 1){
    mtext("Fishing Grounds",
          side = 3,
          line = 1)
  }

  plot(ntr_agg[[i]],
       xlim = c(x_min_global,
                x_max_global),
       ylim = y_lim_group,
       style = "log_y_axis",
       legend_text_second_row_multiplier = 4,
       seg_col = seg_col,
       ylab = "",
       show_fit_on_top = FALSE)

  if(i == 1){
    mtext("No-take Reserve",
          side = 3,
          line = 1)
  }
}
```

## Conclusions from these results

TODO

From theory we would expect a steepening of the size spectrum ($b$ becoming more
negative) under fishing
pressure, because larger organisms get fished out, releasing some predation pressure on
smaller organisms. For a No-take Reserve we would conversely expect a shallowing
of the size spectrum ($b$ become less negative).

Presumably the baseline ecosystem was already experiencing fishing pressure, so
the ideal theoretical results would be the FG staying the same as the baseline
(or maybe shallowing as the area is adjacent to the NTR), and the NTR shallowing.

The table and figure show the following:

Crustacea: From the baseline, FG steepens quite a bit and NTR less so. So NTR
indeed shallower than FG, but it did steepen a little from baseline.

Actinopterygii: From the baseline, FG steepened and NTR got shallower.

Chondrichthyes: From the baseline, FG got shallower, and NTR got even
shallower.

Cephalopoda small: Cephalopoda were tricky to fit, in terms of specifying xmin
and xmax. Strangely, from the baseline the FG got shallower (by quite a bit) and the NTR
steepened, opposite to what we would theoretically expect.

Cephalopoda large: From the baseline, the FG shallowed slightly, while NTR
steepened.

So the first three groups all find the resulting NTR to have a shallower size
spectrum than for the FG, in agreement with the ideal theoretically expected
directions. For both Cephalopoda, the results were opposite to what was
expected, but the data had to be divided into small and large, which could maybe
be done differently.

## Aggregating onto one plot

`aggregate_mlebins()` is used in analysis to combine.

Think I just need to write the plotting function. It should work on
`baseline_agg` etc. which have class `mlebins_list`.

on 7/5/25. Found
some .Rmd code (originally in data-raw/zabala-data/zabala-data-analysis.Rmd but
that is now deleted, copying from GitHub; was the same commit as the new function).

And also `aggregating-functions.R` that contains several functions that are
needed for plotting. Was consistent with sizeSpectra, so just need to update
those for sizeSpectra2. HERE. Copied into `R-aggregating/`, including .Rmd and
.pdf.

Maybe update the pdf first as a new report, as it counts for new methods also
(likely needed as an appendix for the manuscript).





```{r, aggregatefit, eval = FALSE}
x_min_global <- min(res_table_agg$x_min)
x_max_global <- max(res_table_agg$x_max)

# Had this back in May, but saving has changed.


# Already in res_table_agg: b_vec, n_vec, x_min_vec,

# x_max_vec, for simplicity in plotting.
# Save as a mlebins_list object, then create plotting function for that.

# Plot all the data (should be the same way as MLEbins plot. So lots of boxes.
# Could plot the five PLB fits like I did for MLE example.
# Plot the aggregated fit.


```
