---
title: "Summarising the ananlyses of Juliana Zabala's data having wrangled it"
author: "Andrew Edwards and Juliana Zabala"
output: pdf_document
fontsize: 12pt
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  cache = TRUE,
  cache_path = "zabala-summary-cache/",
  fig.path = "zabala-summary-figs-cache/",
  fig.width = 7,
  fig.height = 9
)
  # comment = "#>",
```

```{r, build, echo = FALSE, eval = FALSE}
# To build either run this line or click knit button in RStudio:
rmarkdown::render("zabala-summary.Rmd")
```

```{r, packages, echo = FALSE, cache = FALSE}
library(dplyr)
load_all()
#library(sizeSpectra2)    # Need latest version
```

Copying from `mediterranean-analysis-2/zabala-summary.Rmd`, which I hadn't
updated for `mediterranean-analysis-3/`. Committing before changing anything
further.

Need to have already run `zabala-data-analysis.Rmd` which takes a while to run
and saves three `.rds`
files in the same directory. Or just have the three files. The files contain the
results after some exploration and iterative calculations.

## Results

```{r, loadresults}
baseline_agg <- readRDS("baseline_agg.rds")
fg_agg <- readRDS("fg_agg.rds")
ntr_agg <- readRDS("ntr_agg.rds")
res_table_agg <- readRDS("res_table_agg.rds")
```

Each of the first three is a list for that strata, with: results for each of the five species groups; the
data all aggregated together for a master plot; and a summary table of results
that contains the values needed.

`res_table_agg` is all results regarding $b$ in
one table for simplicity (the same values are parsed out in each strata).

\clearpage

```{r, aggregatetable}
knitr::kable(res_table_agg,
             digits = 2)
```

First, need a global range for the body sizes:
```{r, globalx}
x_min_global <- min(res_table_agg$x_min)
x_max_global <- max(res_table_agg$x_max)
groups <- unique(res_table_agg$Group)    # Assume we're keeping everything in
                                        # the same order, no reason not to
```


TODO Plot all the data (should be the same way as MLEbins plot. So lots of boxes.
 Could plot the five PLB fits like I did for MLE example.
 Plot the aggregated fit.

TODO Need to summarise those results in a plot of $b$ with confidence
intervals.

## Figures -- one figure with 12 panels

Could also plot each one in turn with linear y-axis also will probably want in
an Appendix. Do panel plot first.

```{r, panelplot}
par(mfrow = c(5, 3),    # group in each row, strata in each column
    oma = c(0, 3, 3, 0))

for(i in 1:length(groups)){
  plot(baseline_agg[[i]],
       xlim = c(x_min_global,
                x_max_global),
       log_y_axis = "yes",
       legend_text_second_row_multiplier = 4)

  mtext(groups[i],
        side = 2,
        line = 3)

  if(i ==1){
    mtext("Baseline",
          side = 3,
          line = 1)
  }

  plot(fg_agg[[i]],
       xlim = c(x_min_global,
                x_max_global),
       log_y_axis = "yes",
       legend_text_second_row_multiplier = 4)

  if(i ==1){
    mtext("Fishing Grounds",
          side = 3,
          line = 1)
  }

  plot(ntr_agg[[i]],
       xlim = c(x_min_global,
                x_max_global),
       log_y_axis = "yes",
       legend_text_second_row_multiplier = 4)

  if(i ==1){
    mtext("No-take Reserve",
          side = 3,
          line = 1)
  }
}
```

## Conclusions from these results

From theory we would expect a steepening of the size spectrum ($b$ becoming more
negative) under fishing
pressure, because larger organisms get fished out, releasing some predation pressure on
smaller organisms. For a No-take Reserve we would conversely expect a shallowing
of the size spectrum ($b$ become less negative).

Presumably the baseline ecosystem was already experiencing fishing pressure, so
the ideal theoretical results would be the FG staying the same as the baseline
(or maybe shallowing as the area is adjacent to the NTR), and the NTR shallowing.

The table and figure show the following:

Crustacea: From the baseline, FG steepens quite a bit and NTR less so. So NTR
indeed shallower than FG, but it did steepen a little from baseline.

Actinopterygii: From the baseline, FG steepened and NTR got shallower.

Chondrichthyes: From the baseline, FG got shallower, and NTR got even
shallower.

Cephalopoda small: Cephalopoda were tricky to fit, in terms of specifying xmin
and xmax. Strangely, from the baseline the FG got shallower (by quite a bit) and the NTR
steepened, opposite to what we would theoretically expect.

Cephalopoda large: From the baseline, the FG shallowed slightly, while NTR
steepened.

So the first three groups all find the resulting NTR to have a shallower size
spectrum than for the FG, in agreement with the ideal theoretically expected
directions. For both Cephalopoda, the results were opposite to what was
expected, but the data had to be divided into small and large, which could maybe
be done differently.

## Aggregating onto one plot

I think this can be done, with the size spectra 'added' together to give a
community one. Bit of work still to do that.
