---
output: 
  github_document:
    math_method: "default"
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->
<!-- meth_method default looks to render fine on GitHub, but equations aren't -->
<!-- perfect when locally building and viewing the .html -->

```{r, instructions, eval = FALSE, include = FALSE}
# Build with
# load_all()
rmarkdown::render("README.Rmd")
```
<!-- which builds the .html that can be viewed locally (but isn't pushed to GitHub;
GitHub uses README.md to make the page you see on GitHub). See pacea if want to
save figures.
-->

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "60%"
)
library(tibble)
library(badger)
```

# sizeSpectraFit

<!-- badges: start -->
<!-- `r badge_lifecycle("under development", "orange")` connection timed out, -->
<!-- 17/9/25, but extracting what's needed here, just does not link to a website-->
`r badge_custom("lifecycle", "under development", color = "orange")`
[![R-CMD-check](https://github.com/andrew-edwards/sizeSpectraFit/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/andrew-edwards/sizeSpectraFit/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/andrew-edwards/sizeSpectraFit/branch/main/graph/badge.svg)](https://app.codecov.io/gh/andrew-edwards/sizeSpectraFit?branch=main)
![Visitors](https://api.visitorbadge.io/api/visitors?path=https%3A%2F%2Fgithub.com%2Fandrew-edwards%2FsizeSpectraFit&label=VISITORS&countColor=%23263759&style=flat&labelStyle=lower)
<!-- badges: end -->
<!-- see data-raw/ for logo hexSticker code-->

A streamlined R package for fitting size spectra to ecological data

**Under development, but feel free to start using**

Monday 27th April 2026

Want to fit a size spectrum to your data? This package provides a single
function for fitting, and has standardised `plot()` functions to easily display 
results. Plus there is a lot more. 

Our paper 'No-take reserve improves size spectra and community structure of
demersal megafauna in the Northwestern Mediterranean Sea' has just been accepted
by *Global Ecology and Conservation*, and the pre-proof version is
[available
online](https://www.sciencedirect.com/science/article/pii/S2351989426001769) today!

I have been working a lot on this package recently. Since people may want to
start using it with the release of the above paper, today I am finalising the
vignettes to (soon) post published .html version of them. All the help files and some
automated tests are still being done.

If you can't wait, then all functions are working and are mostly tested, so feel free
to download and use the package. But please check back and redownload when there is no 'under
development' badge or warning above. Note that some functions here may be
improved by then (and so you may have to update any code you have written); once
no longer 'under development' I will ensure back compatibility. 

The actual code associated with the above paper is in the
[report/mediterranean](https://github.com/andrew-edwards/sizeSpectraFit/tree/main/report/mediterranean)
folder as documented in its README file 
[READmediterranean.txt](https://github.com/andrew-edwards/sizeSpectraFit/tree/main/report/mediterranean/READmediterranean.txt). 

Below is some background on size spectra plus simple examples of fitting size spectra to
different types of data using the package. For more details see the vignettes, that are rendered
as .html on this GitHub site and listed below in the [vignettes
section](https://github.com/andrew-edwards/sizeSpectraFit/tree/main#vignettes).


## Background

The size spectrum of an ecological community characterizes how a property, such
as abundance or biomass, varies with body size. Size spectra are often used as
ecosystem indicators of marine systems. Past applications have included
groundfish trawl surveys, visual surveys of fish in kelp forests and coral
reefs, sediment samples of benthic invertebrates and satellite remote sensing of
chlorophyll, as well as terrestrial systems. Various methods have been used to
fit size spectra over the past decades. In our [*Methods in Ecology and
Evolution*
paper](http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12641/full) [1] we
tested eight such methods and
recommended the use of maximum likelihood. In our [*Marine Ecology Progress
Series* paper](https://www.int-res.com/abstracts/meps/v636/p19-33/) [2] we extended the likelihood
method to properly account for the bin structure of data. In our new [*Global
Ecology and
Conservation* accepted paper](https://www.sciencedirect.com/science/article/pii/S2351989426001769)
[3], we developed a general theory for aggregating size spectra
together from different species groups that have differing catchabilities,
culminating in the new normalised aggregated size spectrum. The methods in [1]
and [2] have been widely used in applications around the worldwide (I will add
some examples here). 

This package provides code for fitting size spectra to ecological data using maximum
likelihood methods. This is a new user-friendly package written from the ground up rather than
adapting old code into a package, and with a 
focus on applying methods to users' data rather than reproducing results from
our earlier papers. This overcomes some shortcomings of the earlier
[sizeSpectra](https://github.com/andrew-edwards/sizeSpectra) package.
The methods used are described in detail
in [1]-[3]. The papers should be read
to properly understand the package (rather than have technical mathematical
details repeated too much here).

## Quick start to show ease of use

The core function is `fit_size_spectrum()`, which automatically detects what
kind of data you have, and uses the MLE (maximum likelihood estimate) or MLEbin
(maximum likelihood estimate for binned data) appropriately.

### Individual body sizes

For a vector of numerical numbers representing body sizes (e.g. body
masses of fish), for demonstration we use the simulated values from [1] (which are
saved in the package as `sim_vec`):
```{r, vecfit}
res <- fit_size_spectrum(sim_vec)

plot(res)
```

This matches the recommended Figure 2a of [1], showing the total counts of
individuals $\geq$ each body mass on the x-axis as points, with the red curve
being the fitted power-law distribution and the dashed curves the distribution
using the 95% confidence limits of the size spectra exponent $b$; the MLE of $b$
and the sample
size $n$ are shown in the top-right corner. The list `res` contains
full details for the results, where `b_mle` is the maximum likelihood estimate
of $b$, `b_conf` is the 95% confidence interval, `x` is the original data used
(i.e. `sim_vec` above; only the first 10 values are printed for brevity),
`x_min` and `x_max` are the minimum and maximum possible values of body mass,
and `method` gives the method used for fitting:
```{r, resshow}
res
```

### Body sizes that are binned

Often data are already binned, and only available to a certain resolution
(e.g. 1-2 g, 2-3 g, 3-4 g, etc.). We have the example `sim_vec_binned` saved as
a dataframe (actually a tibble) in the package. We can simply use the same `fit_size_spectrum()`
function as above. The MLEbin method gets used because `sim_vec_binned` is a
dataframe with the correctly defined columns (namely `bin_min`, `bin_max`, and
`bin_count`, representing the minimum and maximum endpoints of each bin and the
count of individuals in that bin, with one row for each bin):

```{r, binfit}
sim_vec_binned

res_mlebin <- fit_size_spectrum(sim_vec_binned)

plot(res_mlebin)
```

Each rectangle represents a bin, with the horizontal span indicating the range
of body masses for that bin, and the vertical span showing the possible range of
the number of individuals with body mass ≥ the body mass of individuals in that
bin (the uncertainty arises because individuals in a bin could really be of any
body mass within the bin).

Again, the full results are saved as a list, with `data` representing the
original data:
```{r, binres}
res_mlebin
```

### Body masses calculated from length data

In [[2]](https://www.int-res.com/abstracts/meps/v636/p19-33/) we introduced the MLEbins method for fitting size spectra while
accounting for species-specific body-mass bins. Unlike the above example, the
bins can overlap. Here we demonstrate our fitting and plotting code for an example data set from
[[3]](https://www.sciencedirect.com/science/article/pii/S2351989426001769). See the paper and the
[fit-data-mlebins.html](http://htmlpreview.github.io/?https://github.com/andrew-edwards/sizeSpectraFit/blob/main/vignettes/fit-data-mlebins.html) 
vignette for full details.

Specifically we fit the size spectrum for the small cephalopoda species group in the
fishing grounds strata (reproducing Fig. B.14 of [[3]](https://www.sciencedirect.com/science/article/pii/S2351989426001769)). For brevity the
wrangling for that (see the
[fit-data-mlebins.html](http://htmlpreview.github.io/?https://github.com/andrew-edwards/sizeSpectraFit/blob/main/vignettes/fit-data-mlebins.html) 
vignette) has already been done and the required dataset saved in the 
package as a tibble of data where each row represents a species-bin combination, including
the count for that bin:
```{r, dataceph}
data_cephsmall_fg
```
Bins represent body mass. Note the overlapping bins in rows 7 and 8, and also in
rows 8 and 9, because
the bins correspond to different species which have different length-weight
coefficients (see [[3]](https://www.sciencedirect.com/science/article/pii/S2351989426001769)).

The minimum body size to
fit, $x_{min}$, is determined using a method described in
[[3]](https://www.sciencedirect.com/science/article/pii/S2351989426001769). Basically,
since the size spectrum is a decreasing distribution, if we have no *a priori*
reason for setting $x_{min}$ then we determine it from the data based on the
peak of a simple histogram. This is
incorporated into the function `determine_xmin_and_fit_mlebins()`. 
```{r, resceph, eval = FALSE}
res_cephsmall_fg <- determine_xmin_and_fit_mlebins(data_cephsmall_fg)  # takes a few minutes
```
<!-- `res_cephsmall_fg` is actually saved as a data object in the package so -->
<!-- this file renders quickly; hence the `eval=FALSE` in the above chunk -->

We can then plot the results: 
```{r, plotceph, fig.height = 12}
plot(res_cephsmall_fg)
```

The top panel is a histogram of total counts of minimum body sizes, using 1-g
body-size bins, used to determine $x_{min}$. The left-most red bin
is the modal bin, and $x_{xmin}$ for this group-strata combination is the minimum
value within that 1-g bin.

Middle and bottom panels (as suggested in Figure 7 of [[2]](https://www.int-res.com/abstracts/meps/v636/p19-33/)) are the fit of the
bounded power-law distribution using the MLEbins method with linear
and logarithmic y-axes, respectively. For each body-mass bin, the green
horizontal line show the range of body sizes, with its value on the
y-axis corresponding to the total number of individuals in bins whose
minima are $\geq$ the bin's minimum; the green lines help to
distinguish each bin when bins are overlapping. The vertical span of
each grey rectangle shows the possible range of number of individuals
with body masses $\geq$ body mass of individuals in that bin
(horizontal span matches the green line). Red curves are fits of the MLE, with dashed lines
showing 95\% confidence intervals.

For full explanations see the vignettes; note that the above fits are very good
because the data have been simulated from a bounded power-law distribution. 

## Vignettes

Three vignettes for sizeSpectraFit go into more details than shown above. Also
see the help files for the functions for more details. The vignettes are
rendered and viewable on GitHub at (TODO please wait for another day or two!):

* [fit-data.html](http://htmlpreview.github.io/?https://github.com/andrew-edwards/sizeSpectraFit/blob/main/vignettes/fit-data.html)
  -- the MLE and MLEbin methods in more detail than shown above, plus with extra
  plotting ideas
* [fit-data-mlebins.html](http://htmlpreview.github.io/?https://github.com/andrew-edwards/sizeSpectraFit/blob/main/vignettes/fit-data-mlebins.html)
  -- the MLEbins method including further results, calculations, explanations, and figure options.
* [fit-aggregate.html](http://htmlpreview.github.io/?https://github.com/andrew-edwards/sizeSpectraFit/blob/main/vignettes/fit-aggregate.html)
  -- our new normalised aggregated size spectrum, as described in
  [[3]](https://www.sciencedirect.com/science/article/pii/S2351989426001769) 

To run and adapt the code yourself, simply download the raw R Markdown files
from
[fit-data.Rmd](https://github.com/andrew-edwards/sizeSpectraFit/blob/main/vignettes/fit-data.Rmd),
[fit-data-mlebins.Rmd](https://github.com/andrew-edwards/sizeSpectraFit/blob/main/vignettes/fit-data-mlebins),
or
[fit-aggregate.Rmd](https://github.com/andrew-edwards/sizeSpectraFit/blob/main/vignettes/fit-aggregate.Rmd).

Run the file locally, and then adapt it for your own data.

## Installation

To install the latest version just:

```
install.packages("remotes")    # If you do not already have the "remotes" package

remotes::install_github("andrew-edwards/sizeSpectraFit")
```

If you get an error like
```
Error in utils::download.file(....)
```
then the connection may be timing out (this happens to me on our work network). Try

```
options(timeout = 1200)
```
and then try and install again. If you get a different error then post an Issue
or contact <a href="mailto:andrew.edwards@dfo-mpo.gc.ca">Andy</a> for help.

## Citation and troubleshooting

If you use `sizeSpectra` in your work then please cite it, currently as the
following (I hope to write a short manuscript about it soon):

Edwards, A. M. (2026). sizeSpectraFit: A streamlined R package for fitting size spectra to ecological data. R package
  version 1.0.0, <https://github.com/andrew-edwards/sizeSpectraFit>.

I do plan to improve, extend, and maintain the package in the future; knowing
it's being used gives extra motivation. 

If you want a citation for LaTeX and R Markdown
bibliographies then after installation type
`citation("sizeSpectraFit")` 
which also gives the GitHub commit from when you installed the package (TODO
coming soon.

Please report any bugs or suggestions as [an
Issue](https://github.com/andrew-edwards/sizeSpectraFit/issues), or
email <a href="mailto:andrew.edwards@dfo-mpo.gc.ca">Andrew Edwards</a> for help.

## References

[1] **Testing and recommending methods for fitting size spectra to data** by Andrew M. Edwards, James P. W. Robinson, Michael J. Plank, Julia K. Baum and Julia L. Blanchard. ***Methods in Ecology and Evolution*** (2017, 8:57-67). Freely available at <http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12641/full>.

[2] **Accounting for the bin structure of data removes bias when fitting size spectra** by Andrew M. Edwards, James P. W. Robinson, Julia L. Blanchard, Julia K. Baum and Michael J. Plank. ***Marine Ecology Progress Series*** (2020, 636:19-33). Freely available at <https://www.int-res.com/abstracts/meps/v636/p19-33/>.

[3] **No-take reserve improves size spectra and community structure of demersal
megafauna in the Northwestern Mediterranean Sea** by 
Juliana Quevedo, Nixon Bahamon, Andrew M. Edwards, Maria Vigo, Pablo Couve,
Jacopo Aguzzi, Jorge Paramo, and Joan B. Company. ***Global Ecology and
Conservation*** [In press](https://www.sciencedirect.com/science/article/pii/S2351989426001769).



<!-- 
### Might still want to copy these dieas from hdiAnalysis:

Could be useful template:

* [results.html](http://htmlpreview.github.io/?https://github.com/andrew-edwards/hdiAnalysis/blob/main/vignettes/results.html)
  -- Designed as a template for users to analyse their own data, by reproducing the results in
  the manuscript as an example.
* [results-extra.html](http://htmlpreview.github.io/?https://github.com/andrew-edwards/hdiAnalysis/blob/main/vignettes/results-extra.html)
  -- Includes further results, calculations, explanations, and figure options.

To run and adapt the code yourself, simply download the raw R Markdown files
from
[results.Rmd](https://github.com/andrew-edwards/hdiAnalysis/blob/main/vignettes/results.Rmd)
or [results-extra.Rmd](https://github.com/andrew-edwards/hdiAnalysis/blob/main/vignettes/results-extra.Rmd).
Run the file locally, and then adapt it for your own data.
-->
