---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->

```{r, instructions, eval = FALSE, include = FALSE}
# Build with
# load_all()
rmarkdown::render("README.Rmd")
```
<!-- which builds the .html that can be viewed locally (but isn't pushed to GitHub;
GitHub uses README.md to make the page you see on GitHub). See pacea if want to
save figures.
-->

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "60%"
)
library(tibble)
library(badger)
```

# sizeSpectraFit

<!-- badges: start -->
<!-- `r badge_lifecycle("under development", "orange")` connection timed out, -->
<!-- 17/9/25, but extracting what's needed here, just does not link to a website-->
`r badge_custom("lifecycle", "under development", color = "orange")`
[![R-CMD-check](https://github.com/andrew-edwards/sizeSpectraFit/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/andrew-edwards/sizeSpectraFit/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/andrew-edwards/sizeSpectraFit/branch/main/graph/badge.svg)](https://app.codecov.io/gh/andrew-edwards/sizeSpectraFit?branch=main)
![Visitors](https://api.visitorbadge.io/api/visitors?path=https%3A%2F%2Fgithub.com%2Fandrew-edwards%2FsizeSpectraFit&label=VISITORS&countColor=%23263759&style=flat&labelStyle=lower)
<!-- badges: end -->
<!-- see data-raw/ for logo hexSticker code-->

A streamlined R package for fitting size spectra to ecological data

**Under development **

All functions should work (they just might change slightly before the formal
release of version 1.0.0). I need to write the vignettes and tests so that users
know what can be done.

For code associated with our manuscript 'No-take
marine reserve improves size spectra and community structure of demersal
megafauna in the Northwestern Mediterranean Sea', see the
[report/mediterranean](https://github.com/andrew-edwards/sizeSpectraFit/tree/main/report/mediterranean)
folder and it's README file 
[READmediterranean.txt](https://github.com/andrew-edwards/sizeSpectraFit/tree/main/report/mediterranean/READmediterranean.txt).

## Background

The size spectrum of an ecological community characterizes how a property, such
as abundance or biomass, varies with body size. Size spectra are often used as
ecosystem indicators of marine systems. Past applications have included
groundfish trawl surveys, visual surveys of fish in kelp forests and coral
reefs, sediment samples of benthic invertebrates and satellite remote sensing of
chlorophyll, as well as terrestrial systems. Various methods have been used to
fit size spectra over the past decades. In our *Methods in Ecology and
Evolution* paper
[[1]](http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12641/full) we tested
eight such methods and
recommended the use of maximum likelihood. In our *Marine Ecology Progress
Series* paper 
[[2]](https://www.int-res.com/abstracts/meps/v636/p19-33/) we extended the likelihood
method to properly account for the bin structure of data. 

This package provides code for fitting size spectra to ecological data using maximum
likelihood methods. This is a new user-friendly package written from the ground up rather than
adapting old code into a package, and with a 
focus on applying methods to users' data rather than reproducing results from
our earlier papers. This overcomes some shortcomings of the earlier
[sizeSpectra](https://github.com/andrew-edwards/sizeSpectra) package.
The methods used are described and fully simulation tested
in [1] and [2]. The papers should be read
to properly understand the package (we aim not to repeat too much from the
papers in the package).


## Fit a vector of values

For a vector of numerical numbers, `x`, representing body sizes (e.g. body
masses of fish), you can simply use the new `fit_size_spectrum()` function. To
demonstrate, we will set `x` to be the simulated values used in [1] (which is
saved in the package as `sim_vec`):
```{r, vecfit}
x <- sim_vec             # replace this with your own data
res <- fit_size_spectrum(x)

plot(res)
```

This matches the recommended Figure 2a of [1].

To fit the data over a restricted range, say between 5 and 200:
```{r, vecfit2}
res_restricted <- fit_size_spectrum(x,
                                    x_min = 5,
                                    x_max = 200)

plot(res_restricted)
```

For more details see the vignettes (not submitted rendered versions yet, so stay
tuned).

<!-- ## Fit a dataframe of values representing binned data -->

## Installation

To install the latest version just:

```
install.packages("remotes")    # If you do not already have the "remotes" package

remotes::install_github("andrew-edwards/sizeSpectraFit")
```

If you get an error like
```
Error in utils::download.file(....)
```
then the connection may be timing out (this happens to me on our work network). Try

```
options(timeout = 1200)
```
and then try and install again. If you get a different error then post an Issue
or contact <a href="mailto:andrew.edwards@dfo-mpo.gc.ca">Andy</a> for help.

### References

[1] **Testing and recommending methods for fitting size spectra to data** by Andrew M. Edwards, James P. W. Robinson, Michael J. Plank, Julia K. Baum and Julia L. Blanchard. ***Methods in Ecology and Evolution*** (2017, 8:57-67). Freely available at <http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12641/full>.

[2] **Accounting for the bin structure of data removes bias when fitting size spectra** by Andrew M. Edwards, James P. W. Robinson, Julia L. Blanchard, Julia K. Baum and Michael J. Plank. ***Marine Ecology Progress Series*** (2020, 636:19-33). Freely available at <https://www.int-res.com/abstracts/meps/v636/p19-33/>.

### Ignore this, still adapting for possible use

Could be useful template:

* [results.html](http://htmlpreview.github.io/?https://github.com/andrew-edwards/hdiAnalysis/blob/main/vignettes/results.html)
  -- Designed as a template for users to analyse their own data, by reproducing the results in
  the manuscript as an example.
* [results-extra.html](http://htmlpreview.github.io/?https://github.com/andrew-edwards/hdiAnalysis/blob/main/vignettes/results-extra.html)
  -- Includes further results, calculations, explanations, and figure options.

To run and adapt the code yourself, simply download the raw R Markdown files
from
[results.Rmd](https://github.com/andrew-edwards/hdiAnalysis/blob/main/vignettes/results.Rmd)
or [results-extra.Rmd](https://github.com/andrew-edwards/hdiAnalysis/blob/main/vignettes/results-extra.Rmd).
Run the file locally, and then adapt it for your own data.
