---
title: "Putting Juliana Zabala's Mediterranean data into sizeSpectra2"
author: "Andrew Edwards and Juliana Zabala"
output: pdf_document
fontsize: 12pt
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  cache = FALSE,
  # cache_path = "zabala-data-cache/",
  # fig.path = "zabala-data-figs-cache/",
  fig.width = 7,
  fig.height = 6
)
  # comment = "#>",
```

```{r, build, echo = FALSE, eval = FALSE}
# To build either run this line or click knit button in RStudio:
rmarkdown::render("mediterranean-data.Rmd")
```

```{r, packages, echo = FALSE}
library(dplyr)
load_all()
library(readxl)
```

```{r, notes, echo = FALSE}

# Was originally data-raw/zabala-data/zabala-data.Rmd, moved to here,
#  renamed, and rerun and simplifying

# Notes from R file:

# Adapting this from using-sizeSpectra2.R in size-spectra-applications, and also
#  using that file for the analyses that are being done in report/
# using-sizeSpectra2.R - size spectrum fits using sizeSpectra2 functions.
#  Run from the same directory as 2023-03-22_Dades_RESNEP.xlsx, or setwd to the
#  required directory.
# Starting from analysis-juliana-2024-12-05.R
# which Andy tweaked (just commented out setwd in the end I think) to get
#  SizeSpectra_TaxonomicGroups_AE - x_min_12mm
# working on his machine. Juliana had sent that on 2024-12-05.

# Size spectra - Andy's code
# Only size spectra Analyses - without GADICULUS ARGENTEUS
# Palamós 2017, 2021
# Juliana Quevedo Zabala
# 08 Oct 2023
```

We are going to save the data in one big tibble, plus the length-weight
coefficients in another.

TODO need to document the data objects.

# Data preparation

Juliana's code to load in and define the data.

```{r, datapreparation, warning = FALSE}
# DATA PREPARATION #############################################################
# DATA PREPARATION: FrequenciaDeTalles tab
resnep_raw <- readxl::read_excel("2023-03-22_Dades_RESNEP.xlsx",3)

resnep <- resnep_raw

resnep$Any <- as.character(resnep$Any)

# no cover data
resnep <- subset(resnep, CodiPesca != "20170823_64500_ESP000023428_03")

# exclude vessel: MEDAN in resnep due to different fishing technique which leads to bias in analysis
resnep <- subset(resnep, NomBarca != "MEDAN")

# exclude species whose length-weight relationship overestimate their weight by more then 200%
resnep <- subset(resnep, NomEspecie != "Sepietta spp." &
                   NomEspecie != "Sepietta oweniana" &
                   NomEspecie != "Histoteutis bonelli" &
                   NomEspecie != "Eledone cirrhosa" &
                   NomEspecie != "Macropipus tuberculatus" &
                   NomEspecie != "Squilla mantis" &
                   NomEspecie != "Scaeurgus unicirrhus" &
                   NomEspecie != "Gadiculus argenteus") # Gadiculus because it can be pelagic and amounts to a lot of variability in 2021
delete_CodiPesca <- "20170824_64500_ESP000023428_04"
delete_Name <- "Liocarcinus depurator"
# Delete rows matching the specified ID and Name
resnep <- resnep %>%
  filter(!(CodiPesca == delete_CodiPesca & NomEspecie == delete_Name))

# exclude from resnep21_fg trawls done beside the NTR -- too close
resnep <- subset(resnep, CodiPesca != "20210826_64500_ESP000023428_09" &
                         CodiPesca != "20210824_64500_ESP000023428_01" &
                         CodiPesca != "20210831_64500_ESP000023428_16")

resnep_for_length_weight_coefficients <- resnep    # To then extract what we
                                                   # need below
```

## Not needed

The following is no longer needed, but keeping here to check I've included
everything in the later code.

```{r, resnepnotneeded}
# Make df smaller
resnep <- select(resnep, CodiPesca, NomPort, Data, Any, NumeroPesca, TipusDeFons, ProfunditatMitjana_m,
                 NomEspecie, ClassificacioCaptura, AgrupamentTaxonomic, Categoria, Talla_mm,
                 FrequenciaCalculada, Abundancia_NIndividusTalla_Km2, PesTotalCalculatRelacioTallaPes_g,
                 Biomassa_KgTalla_Km2, AreaEscombrada_km2, TipusMalla, Site)

# Spread datasets: FrequenciaCalculada - extrapolate subsamples and compute individual weights
resnep <- mutate(resnep, PesoIndividual = PesTotalCalculatRelacioTallaPes_g/FrequenciaCalculada)
resnep <- as.data.frame(lapply(resnep, rep, resnep$FrequenciaCalculada))
resnep['frecuencia_final'] = 1 # new column, frecuencias finales = 1

# Exclude sizes smaller than 12mm, as the net size was 12x12mm, we assume organisms <12mm are not well sampled
resnep <- subset(resnep, Talla_mm > 12)

# Andy doing this to create factors, to ask Juliana about some of the columns.
resnep_temp <- tibble::as_tibble(resnep) %>%
  dplyr::mutate_at(vars(CodiPesca,
                        NomPort,
                        Data,
                        Any,
                        TipusDeFons,
                        NomEspecie,
                        ClassificacioCaptura,
                        AgrupamentTaxonomic,
                        Categoria,
                        TipusMalla, # not sure, says logical but also NA's
                        Site),
                   factor)

summary(dplyr::select(resnep_temp,
                      NumeroPesca,
                      Talla_mm,
                      FrequenciaCalculada,
                      AgrupamentTaxonomic,
                      Abundancia_NIndividusTalla_Km2,
                      PesTotalCalculatRelacioTallaPes_g,
                      Biomassa_KgTalla_Km2,
                      AreaEscombrada_km2,
                      PesoIndividual,
                      frecuencia_final))


# separate resnep 2017 and 2021 (NTR & FG),
resnep17 <- subset(resnep, Any != "2021")
resnep21 <- subset(resnep, Any != "2017")
resnep21_ntr <- subset(resnep21, Site == "MPA")
resnep21_fg <- subset(resnep21, Site == "FG")


# FREQUENCY BARPLOTS, ABUNDANCIA VS TALLA PLOTS, ABUNDANCIA VS PESO PLOTS, METADATA ANALYSES,
# Weight and size supperposed histograms - see relationship, DENSITY HISTOGRAMS BY 'AGRUPAMENT TAXONOMIC':
# find in "SizeSpectra_Analyses_v5.R" ***


# SELECT DATASETS BY 'AGRUPAMENT TAXONOMIC'
# "Crustacea"  "Cephalopoda"  "Actinopterygii"  "Chondrichthyes"  "Echinodermata"
# Exclude Echinidermata due to very low organisms. max 10.

# resnep17
resnep17_crustacea <- subset(resnep17, AgrupamentTaxonomic == "Crustacea")
resnep17_cephalopoda <- subset(resnep17, AgrupamentTaxonomic == "Cephalopoda")
resnep17_actinopterygii <- subset(resnep17, AgrupamentTaxonomic == "Actinopterygii")
resnep17_chondrichthyes <- subset(resnep17, AgrupamentTaxonomic == "Chondrichthyes")

# [Now doing the >10 and <10 at the analysis stage]
resnep17_cephalopoda_small <- subset(resnep17_cephalopoda, PesoIndividual < 10)
resnep17_cephalopoda_big <- subset(resnep17_cephalopoda, PesoIndividual > 10)

# resnep21_fg
resnep21_fg_crustacea <- subset(resnep21_fg, AgrupamentTaxonomic == "Crustacea")
resnep21_fg_cephalopoda <- subset(resnep21_fg, AgrupamentTaxonomic == "Cephalopoda")
resnep21_fg_actinopterygii <- subset(resnep21_fg, AgrupamentTaxonomic == "Actinopterygii")
resnep21_fg_chondrichthyes <- subset(resnep21_fg, AgrupamentTaxonomic == "Chondrichthyes")

resnep21_fg_cephalopoda_small <- subset(resnep21_fg_cephalopoda, PesoIndividual < 10)
resnep21_fg_cephalopoda_big <- subset(resnep21_fg_cephalopoda, PesoIndividual > 10)

# resnep21_ntr
resnep21_ntr_crustacea <- subset(resnep21_ntr, AgrupamentTaxonomic == "Crustacea")
resnep21_ntr_cephalopoda <- subset(resnep21_ntr, AgrupamentTaxonomic == "Cephalopoda")
resnep21_ntr_actinopterygii <- subset(resnep21_ntr, AgrupamentTaxonomic == "Actinopterygii")
resnep21_ntr_chondrichthyes <- subset(resnep21_ntr, AgrupamentTaxonomic == "Chondrichthyes")

resnep21_ntr_cephalopoda_small <- subset(resnep21_ntr_cephalopoda, PesoIndividual < 10)
resnep21_ntr_cephalopoda_big <- subset(resnep21_ntr_cephalopoda, PesoIndividual > 10)
```

Now just extracting and then saving the `PesoIndividual` column for each species
group and data set ("baseline", "fg", or "ntr"), as, for example,
`x_community_ntr`. Using the first syllable or so for the longer complex species
group names.

Baseline corresponds to 2017, but don't need to have the
date in the variable name. Doing the analyses in
`report/`, here just saving the data into the package. Each is then just a
vector of individual body masses. Andy adding in `as.numeric()` as just using
`na.omit()` keeps information (as attributes) about the indices (I presume) of
the omitted NA's, which we do not need.

```{r extractjustmasses}
x_community_baseline <- na.omit(resnep17$PesoIndividual) %>% as.numeric()
x_community_fg <- na.omit(resnep21_fg$PesoIndividual) %>% as.numeric()
x_community_ntr <- na.omit(resnep21_ntr$PesoIndividual) %>% as.numeric()


# Chondrichthyes  [not doing this now in chunks below even though it was here]
resnep17_chondrichthyes <- subset(resnep17_chondrichthyes, Talla_mm >100)
x_chondr_baseline <- na.omit(resnep17_chondrichthyes$PesoIndividual) %>% as.numeric()

x_chondr_fg <- na.omit(resnep21_fg_chondrichthyes$PesoIndividual) %>% as.numeric()

x_chondr_ntr <- na.omit(resnep21_ntr_chondrichthyes$PesoIndividual) %>% as.numeric()

# Cephalopoda -- All
x_ceph_all_baseline <- na.omit(resnep17_cephalopoda$PesoIndividual) %>% as.numeric()
x_ceph_all_fg <- na.omit(resnep21_fg_cephalopoda$PesoIndividual) %>% as.numeric()
x_ceph_all_ntr <- na.omit(resnep21_ntr_cephalopoda$PesoIndividual) %>% as.numeric()
```

Carrying on...


```{r continue}
# Cephalopoda -- Large
x_ceph_large_baseline <- na.omit(resnep17_cephalopoda_big$PesoIndividual) %>% as.numeric()
x_ceph_large_fg <- na.omit(resnep21_fg_cephalopoda_big$PesoIndividual) %>% as.numeric()
x_ceph_large_ntr <- na.omit(resnep21_ntr_cephalopoda_big$PesoIndividual) %>% as.numeric()

# Cephalopoda -- Small
x_ceph_small_baseline <- na.omit(resnep17_cephalopoda_small$PesoIndividual) %>% as.numeric()
x_ceph_small_fg <- na.omit(resnep21_fg_cephalopoda_small$PesoIndividual) %>% as.numeric()
x_ceph_small_ntr <- na.omit(resnep21_ntr_cephalopoda_small$PesoIndividual) %>% as.numeric()

# Actinopterygii
resnep17_actinopterygii <- subset(resnep17_actinopterygii, Talla_mm > 40)
x_actin_baseline <- na.omit(resnep17_actinopterygii$PesoIndividual) %>% as.numeric()

resnep21_fg_actinopterygii <- subset(resnep21_fg_actinopterygii, Talla_mm > 40)
x_actin_fg <- na.omit(resnep21_fg_actinopterygii$PesoIndividual) %>% as.numeric()

resnep21_ntr_actinopterygii <- subset(resnep21_ntr_actinopterygii, Talla_mm >40)
x_actin_ntr <- na.omit(resnep21_ntr_actinopterygii$PesoIndividual) %>% as.numeric()

# Crustacea
x_crust_baseline <- na.omit(resnep17_crustacea$PesoIndividual) %>% as.numeric()
x_crust_fg <- na.omit(resnep21_fg_crustacea$PesoIndividual) %>% as.numeric()
x_crust_ntr <- na.omit(resnep21_ntr_crustacea$PesoIndividual) %>% as.numeric()
```

Do not actually need this, as this was assuming exact measurements for MLE
method, which we are not using. Not running:

```{r, savedata, eval = FALSE}
# This is alphabetical (from using ls()):
mediterranean_megafauna <- tibble::lst(x_actin_baseline,
                                       x_actin_fg,
                                       x_actin_ntr,
                                       x_ceph_all_baseline,
                                       x_ceph_all_fg,
                                       x_ceph_all_ntr,
                                       x_ceph_large_baseline,
                                       x_ceph_large_fg,
                                       x_ceph_large_ntr,
                                       x_ceph_small_baseline,
                                       x_ceph_small_fg,
                                       x_ceph_small_ntr,
                                       x_chondr_baseline,
                                       x_chondr_fg,
                                       x_chondr_ntr,
                                       x_community_baseline,
                                       x_community_fg,
                                       x_community_ntr,
                                       x_crust_baseline,
                                       x_crust_fg,
                                       x_crust_ntr)

usethis::use_data(mediterranean_megafauna,
                  overwrite = TRUE)
```


# For MLEbins method, need tibble of species-specific length-weight coefficients

Juliana already did the length-to-weight calculations in the spreadsheet. But we
need to explicitly do them in the analysis stage as we want the min and max
possible weight for each length bin. So need the coefficients saved.

So, using a version of the data that is saved above:

```{r, lwcoeffs}
length_weight_coefficients <- resnep_for_length_weight_coefficients %>%
  dplyr::select(species = NomEspecie,
                alpha = ConstantA_TallaPes,
                beta = ConstantB_TallaPes) %>%
  dplyr::mutate_at(vars(species),
                   factor)

length_weight_coefficients <-
  dplyr::summarise(dplyr::group_by(length_weight_coefficients,
                                   species),
                   alpha = unique(alpha),
                   beta = unique(beta))

length_weight_coefficients
summary(length_weight_coefficients)

mediterranean_length_weight_coefficients <-
  dplyr::filter(length_weight_coefficients,
                !is.na(alpha))

summary(mediterranean_length_weight_coefficients)

usethis::use_data(mediterranean_length_weight_coefficients,
                  overwrite = TRUE)
# TODO document, alpha and beta are the coeffcicients; need to describe units.
# This should be the standard way of using length-weight coefficients.
```

# For MLEbins method, need the data, somewhat similar to first four columns of Table 1 in MEPS paper

We need the strata and group, species name (not actually needed for analyses but
useful to keep track of), length (with documentation of what that really means),
and number, which is the number of individuals per km2 (alread standardised by
Juliana, confirming calculation below). The number can be non-integer, since it is standardised.

Then in the analysis we will calculate the min and max for each length bin, and
the corresponding min and max for each weight bin based on the species-specific
length-weight coefficients.

Taking the full data set and then save what we need.

Excludings sizes smaller than 12mm, as the net size was 12x12mm, we assume
organisms <12mm are not well sampled. And exclude Echinidermata and Tunicata
groups due to very low number of organisms.

```{r, wrangling}
data_full <- dplyr::filter(resnep_for_length_weight_coefficients,
                           Talla_mm > 12,
                           !(AgrupamentTaxonomic %in% c("Echinodermata",
                                                        "Tunicata")))
                           # Excludes only 21 and 24 rows for each

summary(as.factor(data_full$Site))   # Just need these as strata, TBA is baseline for
                                     # the 4030 in 2017:
nrow(dplyr::filter(data_full,
                   Site == "TBA" & Any == 2017))

# Check the standardisation already done:
expect_equal(data_full$Abundancia_NIndividusTalla_Km2,
             data_full$FrequenciaCalculada / data_full$AreaEscombrada_km2,
             tolerance = 0.0001)
# So I can just use Abundancia_NIndividusTalla_Km2, though would need
# FrequenciaCalculada for MLE method

# Rename to give standard general names for analyses, though likely will
# sometimes have year and no strata or group. But people need to analyse
# strata/year in turn and look at results, so probably not going to fully
# functionalise that step.
data <- dplyr::select(data_full,
                      Site,
                      AgrupamentTaxonomic,
                      NomEspecie,
                      Talla_mm,
                      Abundancia_NIndividusTalla_Km2) %>%
  dplyr::mutate_at(vars(Site,
                        AgrupamentTaxonomic,
                        NomEspecie),
                   factor) %>%
  dplyr::rename(strata = Site,
                group = AgrupamentTaxonomic,
                species = NomEspecie,
                length = Talla_mm,
                number = Abundancia_NIndividusTalla_Km2) %>%
  dplyr::mutate(strata = dplyr::recode(strata,
                                       TBA = "baseline",
                                       FG = "fg",
                                       MPA = "ntr"))  # lower case and be consistent


data

mediterranean_data <- data

summary(mediterranean_data)

usethis::use_data(mediterranean_data,
                  overwrite = TRUE)
```

## Notes from Juliana's emails:

Frequenciecaculada: 3 means measured 1/3 of the catch, or, equivalently (I think)
there were 3 fish but only 1 was measured. So Juliana then creates 3 individual
identical rows in such a situation. 1 means no subsampling.

Each tow covers a different area, we tried to make it so that eachtow last 1
hour, but there were variations, so we use the trawled area km2 to standardize biomass and abundance data.


'PesTotalCalculatRelacioTallaPes_g' is the total weight of that species taking
into account the subsample, meaning that if the total weight is 100g and the
subsample is 5, 5 organisms of that species weight 100g, so each organism here eights 20g. - we won't need that as using length-weight relationships.

NumeroPesca is Haul Number - not needed

FrequenciaCalculada - The number of fish of a given length, species and haul is
represented in FrequenciaCalculada. That's why I did the calculations on the
first question, so that there would be one row for each individual, so using
PesoIndividual accounts for everything.

Abundancia_NIndividusTalla_Km2 this looks to be freq scaled by area of
trawl, as verified in expect_equal above. If you just do MLE you can't
take this into account though.
