---
title: "Fit data and plot results - MLE and MLEbin methods"
author: "Andrew Edwards"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fit data and plot results - MLE and MLEbin methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, instr, echo = FALSE, eval = FALSE}
# To build either run this line or click knit button in RStudio
rmarkdown::render("fit-data.Rmd")
```

```{r, load, include = FALSE}
load_all()   # TODO replace with library(sizeSpectra2)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = " ",
  fig.width = 7,
  fig.height = 6
)
```

Fitting a size spectrum to a data set using likelihood is done using the
function `fit_size_spectrum()`. The function automatically selects the
method to use based on the class of the data, as described by Table 2 in our
[MEPS paper](https://www.int-res.com/abstracts/meps/v636/p19-33/).

TODO For README file. Let's have a simple vector of real measurements, and do
the fitting and plotting functions, without extra explanation. This will show
the simplicity, and refer to vignette for more details. Can use the MLEbins
vignette maybe (when it's done).

## Measurements of individual body sizes -- MLE method

The simplest data is just a vector of measurements, such as body masses or
body lengths. We fit the size spectrum by calculating the maximum likelihood estimate
(the MLE method) of the size-spectrum exponent, along with its 95% confidence intervals.

As an example we use the simulated data from a bounded power-law distribution as used for Figures 1 and 2 of
our [MEE
paper](http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12641/full). The
values are included in sizeSpectra2 as the data object `sim_vec` (which stands
for simulated vector), and we assume they are individual body masses measured in
grammes. Note that the fit will be great
because the data are sampled from the expected power-law distribution.

First, verify that there are 1000 values and print the first 20 values to take a
look:
```{r, MLE1}
length(sim_vec)
sim_vec[1:20]
```

The main fitting function is `fit_size_spectrum()`, which will automatically use
the MLE method here because the data are a vector of individual values:
```{r, MLE2}
res <- fit_size_spectrum(sim_vec)
res
```
The resulting `res` is a list that contains:

- `b_mle`: the MLE of the exponent $b$,
- `b_conf`: the 95% confidence interval of $b$,
- `x`: the original data (the printing here is set up to only show the
 first 10 values),
- `x_min` and `x_max`: the MLEs of $x_{min}$ and $x_{max}$, which are simply the minimum and
 maximum of the data,
- `method`: tells us that the MLE method was used.

The list `res` also has class `size_spectrum_numeric` which helps with our
plotting and printing functions.

To plot the PLB fit and the original data as an Individual Size
Distribution (ISD):
```{r, MLE4}
plot(res)
```
as recommended in Figure 6b of
our [MEE paper](http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12641/full).
Each individual body mass is shown as a point (note the log scales), the solid
red line is the maximum likelihood estimate (MLE) of the size-spectrum exponent
$b$, and the dashed lines are the fits from using the low and high values of the
95% confidence interval of $b$. The MLE for $b$ and the sample size $n$ are
also explicitly given. The default x-axis label assumes the body masses are in
g, but this can easily be changed (see later).

The `style` argument gives other styles of plotting, for example:
```{r, MLElinearplot}
plot(res,
     style = "linear_y_axis")
```
gives the same plot but with a linear y-axis.

To show both the above plots together, with panels automatically labelled as (a)
and (b):
```{r, MLEbothplot}
plot(res,
     style = "both_y_axes")
```
which is similar to Figure 7 of our
[MEPS paper](https://www.int-res.com/abstracts/meps/v636/p19-33/)
but for nonbinned data.

The fourth option is only suitable when the data represent body masses. It
shows the normalised biomass:
```{r, MLEbiomassplot}
plot(res,
     style = "biomass")
```

The black horizontal lines indicate the normalised biomass within each
bin on the y-axis, and the span on the x-axis shows the range of the bin. We use
bars, rather than the usual points (e.g. Figure 6(f) of our MEE
paper) to emphasise that the actual values of body size in each bin span a
range; the points somewhat imply that the midpoint of the bin is more
relevant.

Successive bins double in size, such that they appear of equal width on the log
scale (e.g. Platt and Denman probably TODO). The normalised biomass within each
one is calcluated as the sum of the biomass in the bin divided by the bin width;
there is no uncertainty in this calculation here because we know the individual
body sizes (this is not so if the data are already binned, see later). The
horizontal red lines show the estimated normalised biomass in each bin based on
the MLE of the exponent $b$, and the pink boxes span the normalised biomass
estimated for all values within the 95% confidence interval of $b$. The straight
lines are also shown for the MLE of $b$ (solid line) and the 95% confidency
interval estimates (dashed lines). Note that the straight lines fit the PLB
distribution, whose definition includes the $x_{min}$ and $x_{max}$ bounds (see
MEE paper equation 1) so the lines are bounded, but the top bin here goes beyond $x_{max}$, even though
there are no actual body masses there (the bin happens to extend out further).

The fit here is clearly very good, but remember this is expected as the data are
simulated from a known PLB distribution and are not noisy, unlike real data.

To show the biomass plot and the ISD plot, which is an improved version of the
suggested plot in Figure 6 of our MEE paper:
```{r, MLEbiomassisdplot}
plot(res,
     style = "biomass_and_isd")
```

TODO add in GoF once done.

## Measurements that are already binned -- MLEbin method

Often, data are only available in binned form, motivating our MLEbin and MLEbins
methods (see our MEPS paper for details).

Saved in sizeSpectra2 is a binned version of the simulated data set used above:
```{r}
sim_vec_binned
```
which has information on counts within each body-mass bin, rather than values of
individual
body masses. For example, the first row represent the first bin, which ranges
from 1-2 g (`bin_min` to `bin_max`), and body masses in
this range were observed for 528 individuals (`bin_count`). The bins here progressively double
in width.

To fit the PLB distribution, we again just use the `fit_size_spectrum()`
function (which automatically detects that we want to use the MLEbin method
because our data are a data frame rather than a simple vector of individual body
masses):
```{r}
res_mlebin <- fit_size_spectrum(sim_vec_binned)
res_mlebin
```

So `res_mlebin` is a list object that contains objects:
- `b_mle`: the MLE of the exponent $b$ as calculated using the MLEbin method,
- `b_conf`: the 95% confidence interval of $b$,
- `data`: the original data as used for the fit, plus some extra calculations
 (TODO define somewhere)
- `x_min` and `x_max`: the MLEs of $x_{min}$ and $x_{max}$, which are simply the
minimum of the lowest bin and maximum of the maximum bin,
- `method`: tells us that the MLEbin method was used.
- `attr(,"class")` tells us the list is also of class `size_spectrum_mlebin`,
which helps with our automated plotting and printing functions.

Our default plot for results from the MLEbin method is simply:
```{r, MLEbinplot}
plot(res_mlebin)
```

As for the MLE results above, the `style` argument changes the style of plot:
```{r, MLEbinlinearplot}
plot(res_mlebin,
     style = "linear_y_axis")
```
gives the same plot but with a linear y-axis.

To show both the above plots together:
```{r, MLEbinbothplot}
plot(res_mlebin,
     style = "both_y_axes")
```
which is similar to, but simpler than, Figure 7 of our
[MEPS paper](https://www.int-res.com/abstracts/meps/v636/p19-33/)
because here we do not have species-specific overlapping body-mass bins.

The fourth option is only suitable when the data represent body masses. It
shows the normalised biomass:
```{r, MLEbinbiomassplot}
plot(res_mlebin,
     style = "biomass")
```
and is interpreted the same way as the earlier plot for the MLE method (but here
the bins are already predefined, because that is how the data are provided).

And to show this plot and the first one as two panels:
```{r, MLEbinbiomassandlogplot}
plot(res_mlebin,
     style = "biomass_and_log")
```
The y-axes happen to have the same numbers (though they mean different things) for this simulated data set, though
this will not generally be the case.

Both figures show that the PLB distribution is an excellent fit for these data
the red curves go through the top-left and bottom-right corners of the grey
boxes. Of course, the data are simulated from a PLB distribution and so should
be a good fit, but these figures show the benefit of the plotting approach
(which may not have been so obvious in our MEPS Figure 7 for data with a complex bin
structure).

All the plots are customisable, for example to add extra tick labels to pretty
up a final plot:
```{r, MLEbinplot2)
plot(res_mlebin,
     x_small_ticks_labels = c(5, 50, 500))
```
See `?plot.size_spectrum_numeric` for details.
TODO and _MLEbin? Don't expect so as help should  link them.


HERE

 D make a general `calc_mle_conf` function based on previous `calcLike` (see
   `fit-size-spectrum.numeric.R` also, and move some ideas from that into
   `calc_mle_conf()`, plus replace that code with `calc_mle_conf()`.). Need to make `prof_like()`.

 D use that in `fit-size-spectrum.numeric.R` instead of current

HERE- see mlebins vignette now, and actually the zabala-4 analsysis first.
Need to ensure zabala stuff running, make sure
it still works; think it needs a bit of tweaking.
 - Add/fix tests
 - back to `fit-size-spectrum.data.frame.R` to carry on. See HERE. Doing an
   MLEbins vignette now, since already done code and functions in report/.

 - carry on tidying up help; getting there
 - Change any negll -> neg_ll.
 - see `summary_mle_table()`
 - Check results with paper.
 - Fix TODO's in each plotting function.

And then also consider that we want to also allow for a tibble that has year or
another heading (maybe allow it to be called strata, or let user add that as an
option - yes, strata = year as the default, or maybe strata = NULL is the default).


our [MEE paper](http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12641/full)

our [MEPS paper](https://www.int-res.com/abstracts/meps/v636/p19-33/)


## Session information

```{r sessioninfo}
sessionInfo()
```
