---
title: "Fitting length data and related species-specific body-mass relationships -- MLEbins method"
author: "Andrew Edwards"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitting length data and related species-specific body-mass relationships -- MLEbins method}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, instr, echo = FALSE, eval = FALSE}
# To build either run this line or click knit button in RStudio
rmarkdown::render("fit-data-mlebins.Rmd")
```

```{r, load, include = FALSE}
# load_all()
library(sizeSpectraFit)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = " ",
  fig.width = 7,
  fig.height = 6
)
```

This vignette for the [sizeSpectraFit R
package](https://github.com/andrew-edwards/sizeSpectraFit?tab=readme-ov-file#sizespectrafit)
is for the MLEbins method, which we introduced in our [*MEPS*
paper](https://www.int-res.com/abstracts/meps/v636/p19-33/).
We use an example data set from the Mediterranean from our new 
[*Global Ecology and
Conservation* paper](https://www.sciencedirect.com/science/article/pii/S2351989426001769).

The MLEbins method calculates the maximum likelihood estimate
of the size-spectrum exponent, $b$, when the data are collected as lengths and
we have species-specific length-weight coefficients. So individual body masses
are not available; if they were we could use the MLE method in the
[fit-data.html](http://htmlpreview.github.io/?https://github.com/andrew-edwards/sizeSpectraFit/blob/main/vignettes/fit-data.html)
vignette. If you have binned body masses but the bins overlap, then you also
need to use MLEbins.

The lengths may already be binned, for example a length bin could be 10-15
cm. And even when measured more accurately, the resolution of the measurements
essentially adds uncertainty to a length. For example, a fish of length 15 cm
measured to the nearest cm has a true length in the range 14.5-15.5 cm, which
then gets converted into a body-mass bin using that species specific
length-weight coefficients. The resulting body-mass bins can overlap each other
(unlike for the MLEbin method), because the length bins of differently shaped
species get converted to different body-mass bins. See the [*MEPS*
paper](https://www.int-res.com/abstracts/meps/v636/p19-33/) paper for
further explanation (rather than it being repeated here, where we focus on
applying the method).

## The raw data

We use real data of measured lengths from experimental trawl surveys conducted in
the Northwestern Mediterranean Sea, as part of the research projects RESNEP and
BITER by the Instituto de Ciencias del Mar (ICM-CSIC), Spain; data courtesy of
Juliana Quevedo, Joan B. Company, Nixon Bahamon, and Jordi Ribera Altimir, as
used in
[Quevedo et al. (2026)](https://www.sciencedirect.com/science/article/pii/S2351989426001769). 

The full raw data are available in sizeSpectraFit as
```{r, fulldata}
mediterranean_data
```
for which each row has `strata` (baseline, ntr: the no-take
reserve four years after its implementation, or fg:the fishing grounds four years
after the reserve's implementation), species group given as `group`, `species`, `number` of
individuals of that species that were measured as length `length` (mm). 
Values of `number` can be non-integer because they
are standardised by the area trawled in each tow; the MLEbins method can explicitly
deal with this.

A summary of the data is
```{r, summaryfulldata}
summary(mediterranean_data)
```

We also have the species-specific
length-weight coefficients $\alpha$ and $\beta$ for each species:
```{r, lengthweight}
mediterranean_length_weight_coefficients
```
where the coefficients are for the equation $w = \alpha_s l^{\beta_s}$
relating weights (g) to lengths (mm) for each species $s$.

The above shows what kind of data are needed to implement the MLEbins method.

## Wrangling the data

We first wrangle the data into the format required for using the MLEbins
method. Similar wrangling would need to be done for users' data to get the
required inputs for the likelihood calculations.

For demonstration purposes we will fit the size spectrum for
just one species group (Cephalopoda) and for one strata, the fishing grounds (fg). So we first filter to obtain just those values:
```{r fgcrustacea}
dat <- dplyr::filter(mediterranean_data,
                     strata == "fg",
                     group == "Cephalopoda")
dat
summary(dat)
```
Thus, `dat` contains data for just one `strata` and one `group`, but multiple
`species`. 

The
lengths are in mm, with integer values measured to nearest mm using standard
lab metric tape.
We convert these length values to minima and maxima of bins, with the
measured lengths representing the midpoints. This keeps track of the uncertainty
in the 'true' length of individuals and is needed for the MLEbins method, which
also allows for other species groups having different resolution.

This conversion is done using `calc_bin_breaks()`, which takes a tibble that
contains a column of measurements and appends minimum and maximum values for
each bin (see `?calc_bin_breaks`):

```{r, binbreaks}
dat_with_breaks <- calc_bin_breaks(dat,
                                   bin_width = 1) %>%
  dplyr::rename(bin_count = number)
dat_with_breaks
```

We then convert data of counts of species in length bins to counts in body-mass bins
calculated using the species-specific length-weight coefficients. This is done
using `length_bins_to_body_mass_bins()`
```{r, binbreaksmass}
dat_joined <-
  length_bins_to_body_mass_bins(dat_with_breaks,
                                mediterranean_length_weight_coefficients,
                                length_data_unit = "mm")
dat_joined
```

We then retain just the columns needed to fit using the MLEbins method for input
into `determine_xmin_and_fit_mlebins()`
and also sum the counts for the same combinations of species and weight bins, which is
done in `mediterranean_for_mlebins()`. That function should work for similarly
structured data (requires `group` and `strata`, but the function can be easily
adapted if needed). Since we repeated the same types of analysis for multiple
species groups and strata combinations, it made sense to write a function.
As an extra step here we also are going to look at only small cephalopods below
20.9 g; see
page 8 of Appendix B of [[Quevedo et
al. (2026)]](https://www.sciencedirect.com/science/article/pii/S2351989426001769)
for details.

```{r, combine}
data_cephsmall_fg <- mediterranean_for_mlebins(dat_joined) %>%
  dplyr::filter(bin_min < 20.9) %>%
  dplyr::arrange(bin_min)
data_cephsmall_fg
```
Note the overlapping bins in rows 7 and 8, and also in rows 8 and 9, because the
bins correspond to different species which have different length-weight
coefficients. Hence the need for the MLEbins method.

We are now ready to fit the data. The `determine_xmin_and_fit_mlebins()`
function calculates a histogram to find the peak from which to start the bounded
power-law fit [(Quevedo et
al.,
2026)](https://www.sciencedirect.com/science/article/pii/S2351989426001769). It
then fits the size spectrum. This can take a few minutes, so I have saved
`res_cephsmall_fg` as a data object in sizeSpectraFit (this next chunk is not
actually run, don't tell anyone...; so do make sure to take out `eval = FALSE` if you use this
code as a template):

```{r, mlebinsfit, eval = FALSE}
res_cephsmall_fg <- determine_xmin_and_fit_mlebins(data_cephsmall_fg)
```
The results are
```{r, results}
res_cephsmall_fg
```
This is a list of two objects (each is a list): `$mlebins_fit` contains the
details of the maximum likelihood fitting similar to the other methods, and $h$ is an object of class
`histogram` that contains details of the
histogram used to determine $x_{min}$.

The object has the class `size_spectrum_mlebins` such that our customised
`plot.size_spectrum_mlebins()` automatically gets used: 
```{r, plotmlebins, fig.height = 12}
plot(res_cephsmall_fg)
```
The top panel is a histogram of total counts of minimum body sizes, using 1-g
body-size bins, used to determine $x_{min}$. The left-most red bin
is the modal bin, and $x_{xmin}$ for this group-strata combination is the minimum
value within that 1-g bin.

Middle and bottom panels (as suggested in Figure 7 of [[2]](https://www.int-res.com/abstracts/meps/v636/p19-33/)) are the fit of the
bounded power-law distribution using the MLEbins method with linear
and logarithmic y-axes, respectively. For each body-mass bin, the green
horizontal line show the range of body sizes, with its value on the
y-axis corresponding to the total number of individuals in bins whose
minima are $\geq$ the bin's minimum; the green lines help to
distinguish each bin when bins are overlapping. The vertical span of
each grey rectangle shows the possible range of number of individuals
with body masses $\geq$ body mass of individuals in that bin
(horizontal span matches the green line). Red curves are fits of the MLE, with dashed lines
showing 95\% confidence intervals.

These figure are admittedly a little complex, but retain the full uncertainty
arising from binned data and conversions through species-specific length-weight
relationships. As can be seen, some of the resulting rectangles are quite large,
and simply using a single point for each would suggest higher accuracy than is
warranted.

Do see the help files for relevant functions, and feel free to email me (or make
an issue) with any questions.

## Session information

```{r sessioninfo}
sessionInfo()
```
