---
title: "Fitting length data and related species-specific body-mass relationships -- MLEbins method"
author: "Andrew Edwards"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitting length data and related species-specific body-mass relationships -- MLEbins method}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, instr, echo = FALSE, eval = FALSE}
# To build either run this line or click knit button in RStudio
rmarkdown::render("fit-data-mlebins.Rmd")
```

```{r, load, include = FALSE}
load_all()   # TODO replace with library(sizeSpectraFit)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = " ",
  fig.width = 7,
  fig.height = 6
)
```

This vignette for the [sizeSpectraFit R
package](https://github.com/andrew-edwards/sizeSpectraFit?tab=readme-ov-file#sizespectrafit)
is for the MLEbins method, which we introduced in our [MEPS
paper](https://www.int-res.com/abstracts/meps/v636/p19-33/).
We use an example data set from the Mediterranean from our new 
[Global Ecology and Conservation](https://www.sciencedirect.com/science/article/pii/S2351989426001769).

The MLEbins method calculates the maximum likelihood estimate
of the size-spectrum exponent, $b$, when the data are collected as lengths and
we have species-specific length-weight coefficients. So individual body-masses
are not available; if they were we could use the MLE or MLEbin method in the
[fit-data.html](http://htmlpreview.github.io/?https://github.com/andrew-edwards/sizeSpectraFit/blob/main/vignettes/fit-data.html)
vignette.

The lengths may already be binned, for example a length bin could be 10-11
cm. And even when measured more accurately, the resolution of the measurements
essentially adds uncertainty to a length. For example, a fish of length 15 cm
measured to the nearest cm has a true length in the range 14.5-15.5 cm, which
then gets converted into a body-mass bin using that species specific
length-weight coefficients. The resulting body-mass bins can overlap each other
(unlike for the MLEbin method), because the length bins of differently shaped
species get converted to different body-mass bins. See the MEPS paper for
further explanation (rather than it being repeated here, where we focus on
applying the method).

We use real data of measured lengths from experimental trawl surveys conducted in
the Northwestern Mediterranean Sea, as part of the research projects RESNEP and
BITER by the Instituto de Ciencias del Mar (ICM-CSIC), Spain; data courtesy of
Juliana Quevedo, Joan B. Company, Nixon Bahamon, and Jordi Ribera Altimir. The full data are
available in sizeSpectraFit as simply
```{r meddata}
mediterranean_data
```
and here we analyse the subset of data

HERE - copy from Mediterranean results data-raw, as simplified down. Or README
Or look at med-summary-4.Rmd

and are being analysed using sizeSpectraFit as part of a separate manuscript
(see `?mediterranean_data` for more details). Here we demonstrate the code on a
subset of the data.

## Wrangling the data

We first wrangle the data into the format required for using the MLEbins
method. Similar wrangling would need to be done for users' data to get the
required inputs for the likelihood calculations.

We will fit the size spectrum for
just one species group (crustacea) and for one strata, the fishing grounds that
are close to a no-take marine reserve. So we first filter to obtain just those values:
```{r fgcrustacea}
dat <- dplyr::filter(mediterranean_data,
                     strata == "fg",
                     group == "Crustacea")
dat
summary(dat)
```
Thus, `dat` contains data for just one `strata` and one `group`, but multiple
`species`. The `number` column is the number of individuals of species `species` that were
measured as length `length`. Values of `number` can be non-integer because they
are standardised by the area trawled in each tow; the MLEbins method can explicitly
deal with this.

The
lengths are in mm, with integer values measured to nearest mm using standard
lab metric tape, and values
with decimals are assumed to be measured to 0.01 mm (measured using vernier
calipers). We convert these length values to minima and maxima of bins, with the
measured lengths representing the midpoints. This keeps track of the uncertainty
in the 'true' length of individuals (which obviously is small for the 0.01 mm
bins, but is needed for the MLEbins method), and progresses the uncertainties
through to the likelihood calculations.

TODO see README.Rmd, which I'm copying from test. This wasn't updated below
for 7 months, so is out of date.
The bin breaks are calculated as:
```{r binbreaks}
# TODO convert this all into calc_max_of_bin():
dat_bin_breaks <- calc_max_of_bin(dat$length,
                                  bin_widths = 0.1)

dat_bin_breaks

dat_all <- cbind(dat,
                 dat_bin_breaks) %>%
  tibble::as_tibble() %>%
  dplyr::rename(bin_count = number,
                length_bin_min = bin_min,
                length_bin_max = bin_max)
dat_all
```

using our `calc_max_of_bin()` function, where the argument `bin_widths`
specifies the possible widths of bins (see its help file for details). Assuming
all integer lenghts were measured to the nearest mm will mis-specify about 1% of
measurements (those that were measured to 0.01 mm accuracy but turned out to be
integers), a minor amount.
So we now have minima and maxima of the length bins for each row, defined as
`length_bin_min` and `length_bin_max`, respectively.

In the package we also provide the species-specific length-weight coefficients:
```{r lwcoeffs}
mediterranean_length_weight_coefficients
```
which assume units of cm (not mm) and grammes.

We now use these for each species, to generate body-mass bins:

New function:
D check results agree

```{r convert}
dat_joined <- length_bins_to_body_mass_bins(dat_all,
                                              mediterranean_length_weight_coefficients,
                                              length_data_unit = "mm")
dat_joined
summary(dat_joined)
```

TODO revise: It is useful for users to retain see all the columns above, but here we just
simplify it down (and rename two columns) to what is needed for the fitting,
namely `bin_count`, `bin_min` and `bin_max` (assumed to represent body masses);
users should do the same for their own data.
looking like:


## Fit using MLEbins

Now fit the data using the MLEbins method. First need to determine $x_{min}$....

maybe just do the next and explain what is done.

Those steps have been combined into a single function:
```{r, mlebins}
res <- determine_xmin_and_fit_mlebins(dat_joined)
res

```

Results when used the 0.01 and 1 bins, compare before deleting
```{r oldres, eval = FALSE
#dat_bin_breaks <- calc_max_of_bin(dat$length,
#                                  bin_widths = c(0.01, 1))

$mlebins_fit
 $b_mle
 [1] -1.92739

$b_conf
 [1] -1.93558 -1.91921

$x_min
[1] 2.00599

$x_max
[1] 63.67162

```


TODO explain



```{r, fig}
plot(res)
```



these might work:

```{r, figlinear}
# class(res_2)
# [1] "determine_xmin_and_fit_mlebins" "list"
plot(res$mlebins_fit,
       style = "linear_y_axis")
```

```{r, figlog}
plot(res$mlebins_fit,
       style = "log_y_axis")
```


With the 0.1 mm bins (instead of 0.01 and 1) we now get lots and lots of tiny
bins.

So prob better off picking another data set that was non crustacea, for which
all bins are 1 mm.

Acti baseline maybe?  HERE Maybe redo the zabala results with correct length
resolutions, and then use one as an example. Quick attempt here on Thurs pm:

Quick try, then do big zabala results file: TODO had not taken out big
organisms, so go back to Zabala pdf, tidy that up (simplify function as
described above), then come back here and use just one example. Looks too fine
to use as an example. HERE. Doing that in desktop 3. Need to carry on with them,
the full community ones took quite a while. THen copy an example to here.

```{r actibaseline}
dat_acti <- dplyr::filter(mediterranean_data,
                          strata == "baseline",
                          group == "Actinopterygii")
dat_acti
summary(dat_acti)

dat_acti_bin_breaks <- calc_max_of_bin(dat_acti$length,
                                       bin_widths = 1)

dat_acti_bin_breaks   # looks funny because of rounding

dat_acti_all <- cbind(dat_acti,
                 dat_acti_bin_breaks) %>%
  tibble::as_tibble() %>%
  dplyr::rename(bin_count = number,
                length_bin_min = bin_min,
                length_bin_max = bin_max)
dat_acti_all

dat_acti_joined_3 <- length_bins_to_body_mass_bins(dat_acti_all,
                                              mediterranean_length_weight_coefficients,
                                              length_data_unit = "mm")
dat_acti_joined_3
summary(dat_acti_joined_3)

res_acti <- determine_xmin_and_fit_mlebins(dat_acti_joined_3)
res_acti

plot(res_acti)

```





TODO put check of column names in `determine_xmin_and_fit_mlebins` or similar;
need those in the `select` chunk above. Think this might be okay now.




```{r, exit}
knitr::knit_exit()
```



## Session information

```{r sessioninfo}
sessionInfo()
```
