---
title: "Fit the normalised aggregated size spectrum and plot results"
author: "Andrew Edwards"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fit the normalised aggregated size spectrum and plot results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, instr, echo = FALSE, eval = FALSE}
# To build either run this line or click knit button in RStudio
rmarkdown::render("fit-aggregated.Rmd")
```

```{r, load, include = FALSE}
load_all()   # TODO replace with library(sizeSpectraFit)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = " ",
  fig.width = 7,
  fig.height = 6
)
```

1. Take from report/aggregated example and med-summary-4.Rmd, as the latter will
   have automated some plotting etc.

2. Don't need to explain details too much as refer to Supp Mat.

We demonstrate the normalised aggregated size spectrum here, as introduced in
TODO -- see TODO for details.

TODO check exopontnes in later examples make sense; have to first fix the
legends in basic plots.

The idea is that we can fit size spectra to different parts of the community,
but it is in appropriate to fit a single size spectrum to the full community
because of differences in sampling protocols between different species
groups. For example, the sampling of zooplankton (e.g. using bongo nets) and
pelagic fish (e.g. using a midwater trawl net) is quite different, and we cannot
simply combine the data for a single size spectrum fit. See the TODO above paper
for which we had five different species groups, for which we fit each group
separately and then produced a combined normalised aggregated size
spectrum. This was done for three different strata (a baseline, a no-take
reserve that was protected from fishing, and fishing grounds), showing TODO copy
page 30 text and Figure B.23 here if accepted.

Here we simulate some data to show the code needed to fit a normalised
aggregated size spectrum (NASS).

TODO

This is from aggregating-size-spectra.Rmd, which then got superceded by the
proper analysis. So use this to simulate data, then switch to functionalised
format I had for actual data:

## Generate simulated data

We first simulate $S$ different samples from different bounded power law (PLB) 
distributions, so the fits should be much better than expected from real noisy
data (but the point is to show how to use the code and for users'
understanding).

Each sample has a different sample sizes, exponent, and
range of body sizes, prescribed as vectors:

```{r simulations}
set.seed(42)

S = 4

# For each sample, prescribe the sample size, exponent b, xmin and xmax.
n_vec <- c(6000, 6000, 1600, 2000)
b_vec_known <- c(-1.09, -2, -3, -4)
xmin_known <- c(0.3, 10, 100, 300)
xmax_known <- c(80, 4000, 1000, 1500)

res_list <- list()                      # To save results

for(s in 1:S){
  x_values <- rPLB(n_vec[s],
                   b = b_vec_known[s],
                   xmin = xmin_known[s],
                   xmax = xmax_known[s])

  res_list[[s]] <- fit_size_spectrum(x_values)    # x_values get included in
  # MLE_res[[s]]
}


# Plot each figure with same global x-axis:
xlim_global <- range(res_list[[1]]$x)
for(s in 2:S){
  xlim_global <- range(xlim_global,
                        range(res_list[[s]]$x))
}
xlim_global
```

Here are the four data sets with the four individually fit PLB distributions (the fits are good as data were simulated from PLB distributions):
```{r plotting, fig.height = 8}
# Use the first S of the default colours from plot_aggregate(), for consistency
#  with later plots:
col_vec <- eval(formals(plot_aggregate)$col_vec)[1:S]
  
par(mfrow = c(min(S, 4), 1),
    mai = c(0.6, 0.5, 0.05, 0.3))

for(s in 1:S){
  plot(res_list[[s]],
       xlim = xlim_global,
       fit_col = col_vec[s])
}
```

## Aggregating distributions

So we have `r S` fitted distributions that we want to aggregate, which we can do
and plot with a single function:

```{r plotagg, fig.cap = "The four fitted PLB's from above are shown on the sames axes using the same colour coding as above. Black circles are the full aggregated data set, and the thick magenta curve is the aggregated PLB distribution, the distribution obtained from aggregating the four fitted PLB's together. The black circles are mostly hidden behind the magenta curve for low values of $x$, and at high values are overlaid by the light blue circles because those represent the only data for high $x$. The fits are very good because the data are simulated from PLB's, and likelihood was used to estimated the $b$ exponents."}
orig_agg_fit <- plot_aggregate(res_list)
```

## Simulate disproportionate removal of larger individuals

To show the functionality of using different strata, we take the above data and
disproportionately remove larger fish within each species group. Just removing
10% of the smallest half of the sizes and 50% of the largest half, within each
species group. So not technically removing just big individuals, but this is
just an example to demonstrate plotting.

```{r, removeinds}
res_fished_list <- list()
set.seed(42)    # for reproducibility
for(s in 1:S){
  ind_sizes <- sort(res_list[[s]]$x)

  # Rremove 10% of small fish and 50% of large within the group.
  num_sizes <- length(ind_sizes)   # Assuming this is even
  indices_to_keep <- c(sample(1:(num_sizes/2),
                              size = 0.90 * num_sizes/2),
                       sample( (num_sizes/2 +1):num_sizes,
                              size = 0.50 * num_sizes/2))
  x_values <- ind_sizes[indices_to_keep]
  
  res_fished_list[[s]] <- fit_size_spectrum(x_values)    # x_values get included in
    # MLE_res[[s]]
}
```

Here are the four data sets with the four individually fit PLB distributions
```{r plotting2, fig.height = 8}
par(mfrow = c(min(S, 4), 1),
    mai = c(0.6, 0.5, 0.05, 0.3))

for(s in 1:S){
  plot(res_fished_list[[s]],
       xlim = xlim_global,
       fit_col = col_vec[s])
}
```

And here is the normalised aggregated size spectrum in magenta:
```{r, fishedagg}
fished_agg_fit <- plot_aggregate(res_fished_list)
```

To compare the original and the `fished' normalised aggregated size spectra,
first combine in a list.
```{r, fishedagglist}
agg_list <- list(orig_agg_fit,
                 fished_agg_fit)
strata <- c("unfished",
            "fished")
```
We can then simply use the `plot_aggregate_fits()` function to plot the results
for the two strata on the same graph:

```{r combinedplotnormfunct, fig.cap = "The two aggregated plots restricted to a common range of body masses and then normalised, to properly compare them.", fig.height = 6}
plot_aggregate_fits(agg_list,
                    strata_names = strata,
                    ylim = c(10^(-4), 1))
```

Since there is no simple size-spectrum exponent for such aggregated fits, and
they have different $x_{min}$ values and sample sizes, we restrict the fits to
values above a common value (namely the maximum of the two aggregated
$x_{min}$ values), and then normalise each aggregated fit by its resulting
sample size, such that each starts at the same point (same body-mass on the
x-axis and at 1 on the y-axis, which now represents the proportion (not number) of values
above each body-mass).

This clearly shows a `steepening' of the size spectrum for the fishing grounds
compared to the baseline -- the counts are dropping off faster under
the `fished' scenario. Note the logarithmic
y-axis, such that apparently minor differences are actually quite large. To
further understanding, we also plot the same figure but with a linear y-axis:

```{r combinedplotnormlinear, fig.height = 6}
plot_aggregate_fits(agg_list,
                    strata_names = strata,
                    log_axes = "x")
```

Again, this shows the fished strata size spectrum declining faster than that of
the unfished strata. Note that both curves have to start at the same point, and
reach 0 on the y-axis; minor changes that we see here are therefore indeed of
importance (because the differences can only be so big).

## More extreme removal example

To further illustrate the NASS idea, we repeat the above simulation experiment
but simply remove 50% of all 
individuals > 100 g. The code is largely the same as above, with changes to the
removal of individuals.

```{r, removeinds2}
res_fished_list_2 <- list()
set.seed(42)    # for reproducibility
for(s in 1:S){
  indiv_sizes <- sort(res_list[[s]]$x)

  # Remove 50% of any individuals > 100 g
  num_sizes <- length(indiv_sizes) 
  indiv_sizes_over_100 <- indiv_sizes[indiv_sizes > 100]
  
  x_values <- c(indiv_sizes[indiv_sizes <= 100],
                sample(indiv_sizes_over_100,
                       size = floor(0.5 * length(indiv_sizes_over_100))))
                
  
#  indices_to_keep <- c(whichindiv_sizessample(1:(num_sizes/2),
#                              size = 0.90 * num_sizes/2),
#                       sample( (num_sizes/2 +1):num_sizes,
#                              size = 0.50 * num_sizes/2))
#  x_values <- indiv_sizes[indices_to_keep]
  
  res_fished_list_2[[s]] <- fit_size_spectrum(x_values)    # x_values get included in
    # MLE_res[[s]]
}
```

Here are the four data sets with the four individually fit PLB distributions
```{r plotting3, fig.height = 8}
par(mfrow = c(min(S, 4), 1),
    mai = c(0.6, 0.5, 0.05, 0.3))

for(s in 1:S){
  plot(res_fished_list_2[[s]],
       xlim = xlim_global,
       fit_col = col_vec[s])
}
```

And here is the normalised aggregated size spectrum in magenta:
```{r, fishedagg2}
fished_agg_fit_2 <- plot_aggregate(res_fished_list_2)
```

To compare the original and the `fished' normalised aggregated size spectra,
first combine in a list.
```{r, fishedagglist2}
agg_list_2 <- list(orig_agg_fit,
                 fished_agg_fit_2)
strata <- c("unfished",
            "fished")
```
We can then simply use the `plot_aggregate_fits()` function to plot the results
for the two strata on the same graph:

```{r combinedplotnormfunct2, fig.cap = "The two aggregated plots restricted to a common range of body masses and then normalised, to properly compare them.", fig.height = 6}
plot_aggregate_fits(agg_list_2,
                    strata_names = strata,
                    ylim = c(10^(-4), 1))
```

Since there is no simple size-spectrum exponent for such aggregated fits, and
they have different $x_{min}$ values and sample sizes, we restrict the fits to
values above a common value (namely the maximum of the two aggregated
$x_{min}$ values), and then normalise each aggregated fit by its resulting
sample size, such that each starts at the same point (same body-mass on the
x-axis and at 1 on the y-axis, which now represents the proportion (not number) of values
above each body-mass).

This clearly shows a `steepening' of the size spectrum for the fishing grounds
compared to the baseline -- the counts are dropping off faster under
the `fished' scenario. Note the logarithmic
y-axis, such that apparently minor differences are actually quite large. To
further understanding, we also plot the same figure but with a linear y-axis:

```{r combinedplotnormlinear2, fig.height = 6}
plot_aggregate_fits(agg_list_2,
                    strata_names = strata,
                    log_axes = "x")
```

Again, this shows the fished strata size spectrum declining faster than that of
the unfished strata. Note that both curves have to start at the same point, and
reach 0 on the y-axis; minor changes that we see here are therefore indeed of
importance (because the differences can only be so big).


## Session information

```{r sessioninfo}
sessionInfo()
```
