---
title: "Analysing Juliana Zabala's data having wrangled it"
author: "Andrew Edwards and Juliana Zabala"
output: pdf_document
fontsize: 12pt
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  cache = TRUE,
  cache_path = "zabala-data-analysis-cache/",
  fig.path = "zabala-data-analysis-figs-cache/",
  fig.width = 7,
  fig.height = 9
)
  # comment = "#>",
```

```{r, build, echo = FALSE, eval = FALSE}
# To build either run this line or click knit button in RStudio:
rmarkdown::render("zabala-data-analysis.Rmd")
```

```{r, packages, echo = FALSE, cache = FALSE}
library(dplyr)
load_all()
#library(sizeSpectra2)    # Need latest version
```

Having saved the data in the package via `zabala-data.Rmd`, do the size spectra
analyses here. TODO then move this to report or vignette folder.

Copy and pasting analysis for a particular species group, then just find and
replacing the group name and the short name. Need to examine results and so
can't just make one giant function. The later groups are probably best to use as
templates as the analyses and code got refined along the way.

## Data
We have data and the length-weight coefficients in a standard form. Had analysed
in `zabala-data.Rmd` but want to reduce and functionalise those steps here.

TODO then use same functions to check
Fung results.

So we have three pieces of information. First is the measurements:
```{r medcalcs}
mediterranean_data
summary(mediterranean_data)
```
which we want to fit the size spectrum for each strata-group
combination. The data can have repeated measurements for the same
strata-group-length combination. We will aggregate such examples at the analysis
stage (could do earlier but best to keep data similar to original format).

Second piece is the definition of lengths. They are in mm, with integers measured to nearest mm and ones
with decimals assumed to be measured to 0.01 mm (assuming this will mis-specify, presumably
1%, a few mm ones,
but that's okay). And the lengths assumed to be midpoints (e.g. 10 mm represents
9.5-10.5 mm).

Juliana: The lengths with 0.01mm accuracy are all measured crustaceans, as
they're typically small and its the standard to measure them with a vernier
caliper. All other species were measured to the 1mm accuracy, as we used a
standard lab metric tape. So if we have 27 its correct to say it is 27.00. With
this in mind, I would define b) 26.5 - 27.5 mm as the bin value for 27.00.

And for the 0.01 mm accuracy it really isn't going to make much difference
exactly how
we define the bin, so say 26.13 is 26.125-26.135 for consistency. (Have to
assume this for `calc_max_of_bin()` and need a weight range for each length).


Thirdly, we have species-specific length-weight coefficients.
```{r lwcoeffs}
mediterranean_length_weight_coefficients
```
which assume units of cm (not mm) and grammes.

## Analyses

Pick one combination first, functionalise what we can, then do each in turn as
do need to think about results and xmin choice (i.e. don't do one giant loop).

1. Create weight bins  TODO make a function that's general, but for now can just
do once to get results for each strata-group combination
```{r onestratagroup}
dat_bin_breaks <- calc_max_of_bin(mediterranean_data$length,
                                  bin_widths = c(0.01, 1))
dat_bin_breaks

dat_all <- cbind(mediterranean_data,
                 dat_bin_breaks) %>%
  tibble::as_tibble() %>%
  dplyr::rename(length_bin_min = bin_min,
                length_bin_max = bin_max)

dat_joined <- dplyr::left_join(dat_all,
                               mediterranean_length_weight_coefficients,
                               by = "species") %>%
  dplyr::mutate(weight_bin_min = alpha * (length_bin_min/10)^beta,
                weight_bin_max = alpha * (length_bin_max/10)^beta) %>%
  # remove species with no length-weight coefficients:
  dplyr::filter(!is.na(alpha)) %>%
  dplyr::mutate(strata = recode(strata,
                                TBA = "baseline",
                                FG = "fg",
                                MPA = "ntr"))       # lower case and be consistent

# TODO Could add argument for 10 depending on mm or cm

dat_joined[1:5, ] %>% a()

# Useful for users to be able to see all this, even though only some of it is
# needed for the likelihood fitting.
```

Just want to analyse one strata at a time. Do single baseline-actin here as an example
to help explain then functionalise and do each in turn.

```{r, analyse}
dat_baseline_actin <- dplyr::filter(dat_joined,
                                    strata == "baseline",
                                    group == "Actinopterygii")

# Now strip down to just what's needed, just need counts and bin definitions for
# weights, but keep species for understandability as am outputting the data in
# results, though am combining results of same species and bin_min (so do need
# species at this stage)...
dat_for_mlebins_pre <- dplyr::select(dat_baseline_actin,
                                     species,
                                     bin_count = number,
                                     bin_min = weight_bin_min,
                                     bin_max = weight_bin_max)

dat_for_mlebins_pre
xx <- table(table(dat_for_mlebins_pre$bin_min))
xx

dat_for_mlebins <- dplyr::summarise(dplyr::group_by(dat_for_mlebins_pre,
                                                    species,
                                                    bin_min,
                                                    bin_max),    # not needed
                                                      # for grouping but need to retain
                                    bin_count = sum(bin_count)) %>%
  dplyr::ungroup()

dat_for_mlebins

expect_equal(sum(xx),             # expected resulting number of rows, based on
                                  # unique counts
             nrow(dat_for_mlebins))


res_mlebins <- fit_size_spectrum(dat_for_mlebins)

res_mlebins
```



```{r plotmlebins}
plot(res_mlebins)   # worked, but hadn't done the groups yet.
```

That plot shows very few small organisms, emphasising the need to determine a
value for xmin.

### Determine xmin

This works for calculating xmin:

```{r xminandfitplot}
fit <- determine_xmin_and_fit_mlebins(dat_for_mlebins)

plot(fit, xlim_hist = c(0, 60))
```

## Functionalise

Repeat the above but write a function for it.

```{r functionalised}

data_baseline_actin <- mediterranean_for_mlebins(dat_joined,
                                                 strata_name = "baseline",
                                                 group_name = "Actinopterygii")

expect_equal(dat_for_mlebins,
             data_baseline_actin)
# So function matches what we did in detail above.

res_baseline_actin <- determine_xmin_and_fit_mlebins(data_baseline_actin)

expect_equal(fit,
             res_baseline_actin)
# And functionalised results match what we did above.
```

Now to plot the results:
```{r baselineactinplot}
plot(res_baseline_actin,
     xlim_hist = c(0, 60))
```

Okay, happy with functionalisation. Now to actually analyse the data sets in
turn. Do in the same order of Juliana's big figure.

Adapting the next sections from mediterranean-analysis.Rmd which used the MLE
method. Shapes of resulting ISD plots should somewhat resemble those that
Juliana produced, though we now have xmin more explicit and are using MLEbins.

Need to do all these manually to check results and think about them, and to
determine on a communal xmin (and maybe xmax ) for each species group.

\clearpage


# Full community

Doing full community just to have a look. Know a priori that this is not
appropriate because of the different sampling approaches on the different
groups. Also, the results show clear distinct ranges of data. Just doing for the
baseline.

## Full Community baseline

```{r, fullbaseline}
data_full_baseline <- mediterranean_for_mlebins(dat_joined,
                                                strata_name = "baseline")

res_full_baseline <- determine_xmin_and_fit_mlebins(data_full_baseline)

res_full_baseline$mlebins_fit      # full list includes histogram values,
                                   # could make a print function TODO
```

```{r, fullbaselineplot}
plot(res_full_baseline,
     xlim_hist = c(0, 60),
     main = "Full Community baseline")
```


\clearpage

# Crustacea

Juliana had xmin as 0.4 for FG and 0.3 for NTR; check what it was for
baseline. But determining explicitly here.

## Crustacea baseline

```{r, crustbaseline}
data_crust_baseline <- mediterranean_for_mlebins(dat_joined,
                                                 group_name = "Crustacea",
                                                 strata_name = "baseline")


res_crust_baseline <- determine_xmin_and_fit_mlebins(data_crust_baseline)

res_crust_baseline$mlebins_fit     # full list includes histogram values, TODO
                                   # could make a print function
```

```{r, crustbaselineplot}
plot(res_crust_baseline,
     xlim_hist = c(0, 60),
     main = "Crustacea baseline")
```

## Crustacea FG

```{r, crustfg}
data_crust_fg <- mediterranean_for_mlebins(dat_joined,
                                           group_name = "Crustacea",
                                           strata_name = "fg")

res_crust_fg <- determine_xmin_and_fit_mlebins(data_crust_fg)

res_crust_fg$mlebins_fit
```

```{r, crustfgplot}
plot(res_crust_fg,
     xlim_hist = c(0, 60),
     main = "Crustacea FG")
```



## Crustacea NTR

```{r, crustntr}
data_crust_ntr <- mediterranean_for_mlebins(dat_joined,
                                            group_name = "Crustacea",
                                            strata_name = "ntr")

res_crust_ntr <- determine_xmin_and_fit_mlebins(data_crust_ntr)

res_crust_ntr$mlebins_fit
```

```{r, crustntrplot}
plot(res_crust_ntr,
     xlim_hist = c(0, 60),
     main = "Crustacea NTR")
```

Now to compare the three strata for Crustacea.

```{r, crustsumm}
res_crust_baseline$mlebins_fit$b_mle
res_crust_fg$mlebins_fit$b_mle
res_crust_ntr$mlebins_fit$b_mle

add_row_to_summary_table <- function(summary, res, group, strata){
  rbind(summary,
        dplyr::tibble("Group" = group,
                      "Strata" = strata,
                      "x_min" = res$mlebins_fit$x_min,
                      "x_max" = res$mlebins_fit$x_max,
                      "Low b" = res$mlebins_fit$b_conf[1],
                      "MLE b" = res$mlebins_fit$b_mle,
                      "High b" = res$mlebins_fit$b_conf[2]))
}

res_table_crust <- dplyr::tibble()
res_table_crust <- add_row_to_summary_table(res_table_crust,
                                      res_crust_baseline,
                                      "Crustacea",
                                      "Baseline")

res_table_crust <- add_row_to_summary_table(res_table_crust,
                                      res_crust_fg,
                                      "Crustacea",
                                      "fg")
res_table_crust <- add_row_to_summary_table(res_table_crust,
                                      res_crust_ntr,
                                      "Crustacea",
                                      "ntr")
```

Results combined are:

```{r, rescrusttable}
knitr::kable(res_table_crust, digits = 2)
```

## Global xmin and xmax for Crustacea

So doing the modal approach gave the same xmin for each strata, which is
great. xmax was set independently, so we should select the globally largest for
this species group.

```{r, crustxmax}
crust_xmin <- max(res_table_crust$x_min)
crust_xmin

crust_xmax <- max(res_table_crust$x_max)
crust_xmax
```
We know that is for NTR, so redo the other two with this xmax value.

## Crustacea baseline with fixed xmin and xmax

```{r, crustbaselinexmax}
res_crust_baseline_fix_xmax <-
  determine_xmin_and_fit_mlebins(data_crust_baseline,
                                 x_min = crust_xmin,
                                 x_max = crust_xmax)
```

```{r, crustbaselineplotxmax}
plot(res_crust_baseline_fix_xmax,
     xlim_hist = c(0, 60),
     main = "Crustacea baseline")
```

## Crustacea FG with fixed xmin and xmax

```{r, crustfgxmax}
res_crust_fg_fix_xmax <- determine_xmin_and_fit_mlebins(data_crust_fg,
                                                        x_min = crust_xmin,
                                                        x_max = crust_xmax)
```

```{r, crustfgplotxmax}
plot(res_crust_fg_fix_xmax,
     xlim_hist = c(0, 60),
     main = "Crustacea FG")
```

## Crustacea NTR with fixed xmin and xmax

```{r, crustbaelinexmax}
res_crust_ntr_fix_xmax <- determine_xmin_and_fit_mlebins(data_crust_ntr,
                                                         x_min = crust_xmin,
                                                         x_max = crust_xmax)
```

```{r, crustntrplotxmax}
plot(res_crust_ntr_fix_xmax,
     xlim_hist = c(0, 60),
     main = "Crustacea NTR")
```

Now to compare the three strata for Crustacea with xmax fixed:

```{r, crustsummfix}
res_table_crust_fix_xmax <- dplyr::tibble()
res_table_crust_fix_xmax <- add_row_to_summary_table(res_table_crust_fix_xmax,
                                      res_crust_baseline_fix_xmax,
                                      "Crustacea",
                                      "Baseline")

res_table_crust_fix_xmax <- add_row_to_summary_table(res_table_crust_fix_xmax,
                                      res_crust_fg_fix_xmax,
                                      "Crustacea",
                                      "fg")
res_table_crust_fix_xmax <- add_row_to_summary_table(res_table_crust_fix_xmax,
                                      res_crust_ntr_fix_xmax,
                                      "Crustacea",
                                      "ntr")
```

Results combined are:

```{r, rescrusttablexmax}
knitr::kable(res_table_crust_fix_xmax,
             digits = 2)
```

## Further communities

For each one, copy the above, find and replace relevant community name, but then
manually determine the xmax.


# Actinopterygii

Juliana removed lengths <100mm, so doing that here.

## Actinopterygii baseline

```{r, actinbaseline}
data_actin_baseline <- mediterranean_for_mlebins(dat_joined,
                                                 group_name = "Actinopterygii",
                                                 strata_name = "baseline",
                                                 minimum_length = 100)


res_actin_baseline <- determine_xmin_and_fit_mlebins(data_actin_baseline)

res_actin_baseline$mlebins_fit     # full list includes histogram values, TODO
                                   # could make a print function
```

```{r, actinbaselineplot}
plot(res_actin_baseline,
     xlim_hist = c(0, 60),
     main = "Actinopterygii baseline")
```

## Actinopterygii FG

```{r, actinfg}
data_actin_fg <- mediterranean_for_mlebins(dat_joined,
                                           group_name = "Actinopterygii",
                                           strata_name = "fg",
                                           minimum_length = 100)

res_actin_fg <- determine_xmin_and_fit_mlebins(data_actin_fg)

res_actin_fg$mlebins_fit
```



```{r, actinfgplot}
plot(res_actin_fg,
     xlim_hist = c(0, 60),
     main = "Actinopterygii FG")
```



## Actinopterygii NTR

```{r, actinntr}
data_actin_ntr <- mediterranean_for_mlebins(dat_joined,
                                            group_name = "Actinopterygii",
                                            strata_name = "ntr",
                                            minimum_length = 100)

res_actin_ntr <- determine_xmin_and_fit_mlebins(data_actin_ntr)

res_actin_ntr$mlebins_fit
```



```{r, actinntrplot}
plot(res_actin_ntr,
     xlim_hist = c(0, 60),
     main = "Actinopterygii NTR")
```

Now to compare the three strata for Actinopterygii.

```{r, actinsumm}
res_table_actin <- dplyr::tibble()
res_table_actin <- add_row_to_summary_table(res_table_actin,
                                      res_actin_baseline,
                                      "Actinopterygii",
                                      "Baseline")

res_table_actin <- add_row_to_summary_table(res_table_actin,
                                      res_actin_fg,
                                      "Actinopterygii",
                                      "fg")
res_table_actin <- add_row_to_summary_table(res_table_actin,
                                      res_actin_ntr,
                                      "Actinopterygii",
                                      "ntr")
```

Results combined are:

```{r, resactintable}
knitr::kable(res_table_actin, digits = 2)
```



## Global xmin and xmax for Actinopterygii

So doing the modal approach gave different xmin between baseline and other two;
should go with the largest of those. Though might give quite shallow power law
for baseline.

Alternative is to use xmin for the aggregated data, TODO try that also.

xmax was set independently, so we should select the globally largest for
this species group.

```{r, actinxmax}
actin_xmin <- max(res_table_actin$x_min)
actin_xmin
actin_xmax <- max(res_table_actin$x_max)
actin_xmax
```

## Actinopterygii baseline with fixed xmin and xmax

```{r, actinbaselinexmax}
res_actin_baseline_fix_xmax <-
  determine_xmin_and_fit_mlebins(data_actin_baseline,
                                 x_min = actin_xmin,
                                 x_max = actin_xmax)
```



```{r, actinbaselineplotxmax}
plot(res_actin_baseline_fix_xmax,
     xlim_hist = c(0, 60),
     main = "Actinopterygii baseline")
```

## Actinopterygii FG with fixed xmin and xmax

```{r, actinfgxmax}
res_actin_fg_fix_xmax <- determine_xmin_and_fit_mlebins(data_actin_fg,
                                                        x_min = actin_xmin,
                                                        x_max = actin_xmax)
```



```{r, actinfgplotxmax}
plot(res_actin_fg_fix_xmax,
     xlim_hist = c(0, 60),
     main = "Actinopterygii FG")
```

## Actinopterygii NTR with fixed xmin and xmax

```{r, actinbaelinexmax}
res_actin_ntr_fix_xmax <- determine_xmin_and_fit_mlebins(data_actin_ntr,
                                                         x_min = actin_xmin,
                                                         x_max = actin_xmax)
```

```{r, actinntrplotxmax}
plot(res_actin_ntr,
     xlim_hist = c(0, 60),
     main = "Actinopterygii NTR")
```

Now to compare the three strata for Actinopterygii with xmax fixed:

```{r, actinsummfix}
res_table_actin_fix_xmax <- dplyr::tibble()
res_table_actin_fix_xmax <- add_row_to_summary_table(res_table_actin_fix_xmax,
                                      res_actin_baseline_fix_xmax,
                                      "Actinopterygii",
                                      "Baseline")

res_table_actin_fix_xmax <- add_row_to_summary_table(res_table_actin_fix_xmax,
                                      res_actin_fg_fix_xmax,
                                      "Actinopterygii",
                                      "fg")
res_table_actin_fix_xmax <- add_row_to_summary_table(res_table_actin_fix_xmax,
                                      res_actin_ntr_fix_xmax,
                                      "Actinopterygii",
                                      "ntr")
```

Results combined are:

```{r, resactintablexmax}
knitr::kable(res_table_actin_fix_xmax, digits = 2)
```

\clearpage

# Chondrichthyes

Juliana removed lengths <100mm, so doing that here.

## Chondrichthyes baseline

```{r, chondrbaseline}
data_chondr_baseline <- mediterranean_for_mlebins(dat_joined,
                                                 group_name = "Chondrichthyes",
                                                 strata_name = "baseline",
                                                 minimum_length = 100)


res_chondr_baseline <- determine_xmin_and_fit_mlebins(data_chondr_baseline)

res_chondr_baseline$mlebins_fit
```

```{r, chondrbaselineplot}
plot(res_chondr_baseline,
     xlim_hist = c(0, 300),
     main = "Chondrichthyes baseline")
```

## Chondrichthyes FG

```{r, chondrfg}
data_chondr_fg <- mediterranean_for_mlebins(dat_joined,
                                           group_name = "Chondrichthyes",
                                           strata_name = "fg",
                                           minimum_length = 100)

res_chondr_fg <- determine_xmin_and_fit_mlebins(data_chondr_fg)

res_chondr_fg$mlebins_fit
```



```{r, chondrfgplot}
plot(res_chondr_fg,
     xlim_hist = c(0, 300),
     main = "Chondrichthyes FG")
```



## Chondrichthyes NTR

```{r, chondrntr}
data_chondr_ntr <- mediterranean_for_mlebins(dat_joined,
                                            group_name = "Chondrichthyes",
                                            strata_name = "ntr",
                                            minimum_length = 100)

res_chondr_ntr <- determine_xmin_and_fit_mlebins(data_chondr_ntr)

res_chondr_ntr$mlebins_fit
```



```{r, chondrntrplot}
plot(res_chondr_ntr,
     xlim_hist = c(0, 300),
     main = "Chondrichthyes NTR")
```

Now to compare the three strata for Chondrichthyes.

```{r, chondrsumm}
res_table_chondr <- dplyr::tibble()
res_table_chondr <- add_row_to_summary_table(res_table_chondr,
                                      res_chondr_baseline,
                                      "Chondrichthyes",
                                      "Baseline")

res_table_chondr <- add_row_to_summary_table(res_table_chondr,
                                      res_chondr_fg,
                                      "Chondrichthyes",
                                      "fg")
res_table_chondr <- add_row_to_summary_table(res_table_chondr,
                                      res_chondr_ntr,
                                      "Chondrichthyes",
                                      "ntr")
```

Results combined are:

```{r, reschondrtable}
knitr::kable(res_table_chondr, digits = 2)
```



## Global xmin and xmax for Chondrichthyes

So doing the modal approach gave different xmin between baseline and other two;
should go with the largest of those. Though might give quite shallow power law
for baseline.

Alternative is to use xmin for the aggregated data, TODO try that also.

xmax was set independently, so we should select the globally largest for
this species group.

```{r, chondrxmax}
chondr_xmin <- max(res_table_chondr$x_min)
chondr_xmin
chondr_xmax <- max(res_table_chondr$x_max)
chondr_xmax
```

## Chondrichthyes baseline with fixed xmin and xmax

```{r, chondrbaselinexmax}
res_chondr_baseline_fix_xmax <-
  determine_xmin_and_fit_mlebins(data_chondr_baseline,
                                 x_min = chondr_xmin,
                                 x_max = chondr_xmax)
```



```{r, chondrbaselineplotxmax}
plot(res_chondr_baseline_fix_xmax,
     xlim_hist = c(0, 300),
     main = "Chondrichthyes baseline")
```

## Chondrichthyes FG with fixed xmin and xmax

```{r, chondrfgxmax}
res_chondr_fg_fix_xmax <- determine_xmin_and_fit_mlebins(data_chondr_fg,
                                                        x_min = chondr_xmin,
                                                        x_max = chondr_xmax)
```



```{r, chondrfgplotxmax}
plot(res_chondr_fg_fix_xmax,
     xlim_hist = c(0, 300),
     main = "Chondrichthyes FG")
```

## Chondrichthyes NTR with fixed xmin and xmax

```{r, chondrbaelinexmax}
res_chondr_ntr_fix_xmax <- determine_xmin_and_fit_mlebins(data_chondr_ntr,
                                                         x_min = chondr_xmin,
                                                         x_max = chondr_xmax)
```

```{r, chondrntrplotxmax}
plot(res_chondr_ntr_fix_xmax,
     xlim_hist = c(0, 300),
     main = "Chondrichthyes NTR")
```

Now to compare the three strata for Chondrichthyes with xmax fixed:

```{r, chondrsummfix}
res_table_chondr_fix_xmax <- dplyr::tibble()
res_table_chondr_fix_xmax <- add_row_to_summary_table(res_table_chondr_fix_xmax,
                                      res_chondr_baseline_fix_xmax,
                                      "Chondrichthyes",
                                      "Baseline")

res_table_chondr_fix_xmax <- add_row_to_summary_table(res_table_chondr_fix_xmax,
                                      res_chondr_fg_fix_xmax,
                                      "Chondrichthyes",
                                      "fg")
res_table_chondr_fix_xmax <- add_row_to_summary_table(res_table_chondr_fix_xmax,
                                      res_chondr_ntr_fix_xmax,
                                      "Chondrichthyes",
                                      "ntr")
```

Results combined are:

```{r, reschondrtablexmax}
knitr::kable(res_table_chondr_fix_xmax, digits = 2)
```



## Global xmin and xmax for Chondrichthyes -- try higher bin width for xmin since xmin came out very high and data span a huge range

Just do the xmin calculations and plots, then use those to determine
xmin. Don't really help though. But the above-calculated xmin is not
helpful. TODO Ask Juliana about the data.

```{r, chondrlargerbinwidth}
h_chondr_baseline <- make_hist_for_binned_counts(data_chondr_baseline,
                                                 bin_width = 10,
                                                 bin_start = 0)
x_min_chondr_baseline <- determine_xmin(h_chondr_baseline)

x_min_chondr_baseline

plot(h_chondr_baseline)   # Just doesn't have the colours

# fg

h_chondr_fg <- make_hist_for_binned_counts(data_chondr_fg,
                                                 bin_width = 10,
                                                 bin_start = 0)
x_min_chondr_fg <- determine_xmin(h_chondr_fg)

x_min_chondr_fg

plot(h_chondr_fg)   # Just doesn't have the colours

# ntr

h_chondr_ntr <- make_hist_for_binned_counts(data_chondr_ntr,
                                                 bin_width = 10,
                                                 bin_start = 0)
x_min_chondr_ntr <- determine_xmin(h_chondr_ntr)

x_min_chondr_ntr

plot(h_chondr_ntr)   # Just doesn't have the colours
```

## Global xmin and xmax for Chondrichthyes -- try basing on baseline, which had a very clear xmin

If the large individuals are now showing up in NTR in relatively higher numbers,
it's a bit inconsistent to use that as the starting point for xmin.

```{r, chondrxmax2}
chondr_xmin2 <- dplyr::filter(res_table_chondr, Strata == "Baseline")$x_min
chondr_xmin2
chondr_xmax    # same
```

Appending results and code chunk names with 2.

## Chondrichthyes baseline with fixed xmin and xmax

```{r, chondrbaselinexmax2}
res_chondr_baseline_fix_xmax2 <-
  determine_xmin_and_fit_mlebins(data_chondr_baseline,
                                 x_min = chondr_xmin2,
                                 x_max = chondr_xmax)
```



```{r, chondrbaselineplotxmax2}
plot(res_chondr_baseline_fix_xmax2,
     xlim_hist = c(0, 300),
     main = "Chondrichthyes baseline")
```

## Chondrichthyes FG with fixed xmin and xmax

```{r, chondrfgxmax2}
res_chondr_fg_fix_xmax2 <- determine_xmin_and_fit_mlebins(data_chondr_fg,
                                                          x_min = chondr_xmin2,
                                                          x_max = chondr_xmax)
```



```{r, chondrfgplotxmax2}
plot(res_chondr_fg_fix_xmax2,
     xlim_hist = c(0, 300),
     main = "Chondrichthyes FG")
```

## Chondrichthyes NTR with fixed xmin and xmax

```{r, chondrbaelinexmax2}
res_chondr_ntr_fix_xmax2 <- determine_xmin_and_fit_mlebins(data_chondr_ntr,
                                                           x_min = chondr_xmin2,
                                                           x_max = chondr_xmax)
```

```{r, chondrntrplotxmax2}
plot(res_chondr_ntr_fix_xmax2,
     xlim_hist = c(0, 300),
     main = "Chondrichthyes NTR")
```

Now to compare the three strata for Chondrichthyes with xmin2 and xmax fixed:

```{r, chondrsummfix2}
res_table_chondr_fix_xmax2 <- dplyr::tibble()
res_table_chondr_fix_xmax2 <- add_row_to_summary_table(res_table_chondr_fix_xmax2,
                                      res_chondr_baseline_fix_xmax2,
                                      "Chondrichthyes",
                                      "Baseline")

res_table_chondr_fix_xmax2 <- add_row_to_summary_table(res_table_chondr_fix_xmax2,
                                      res_chondr_fg_fix_xmax2,
                                      "Chondrichthyes",
                                      "fg")
res_table_chondr_fix_xmax2 <- add_row_to_summary_table(res_table_chondr_fix_xmax2,
                                      res_chondr_ntr_fix_xmax2,
                                      "Chondrichthyes",
                                      "ntr")
```

Results combined are:

```{r, reschondrtablexmax2}
knitr::kable(res_table_chondr_fix_xmax2, digits = 2)
```

# Cephalopoda (all)

Juliana had xmin of 2 and xmax 200 (TODO check if lengths or weights). For
consistency we should first analyse the full data in one go first. Results for
MLE (not MLEbins) clearly showed the need to split up.

Small was body weights $<10$ and large was $>10$. First do all together, with
xmin and xmax for each and then common xmin and xmax. These clearly show up the
natural disconnect, motivating the need for doing small and large separately, as
Juliana had aleady found.

Then we work out the value to split them, and re-analyse small and large
separately in the usual way.

## Cephalopoda (all) baseline

```{r, cephallbaseline}
data_cephall_baseline <- mediterranean_for_mlebins(dat_joined,
                                                   group_name = "Cephalopoda",
                                                   strata_name = "baseline")

res_cephall_baseline <- determine_xmin_and_fit_mlebins(data_cephall_baseline)

res_cephall_baseline$mlebins_fit
```



```{r, cephallbaselineplot}
plot(res_cephall_baseline,
     xlim_hist = c(0, 60),
     main = "Cephalopoda (all) baseline")
```

## Cephalopoda (all) FG

```{r, cephallfg}
data_cephall_fg <- mediterranean_for_mlebins(dat_joined,
                                             group_name = "Cephalopoda",
                                             strata_name = "fg")

res_cephall_fg <- determine_xmin_and_fit_mlebins(data_cephall_fg)

res_cephall_fg$mlebins_fit
```



```{r, cephallfgplot}
plot(res_cephall_fg,
     xlim_hist = c(0, 60),
     main = "Cephalopoda (all) FG")
```



## Cephalopoda (all) NTR

```{r, cephallntr}
data_cephall_ntr <- mediterranean_for_mlebins(dat_joined,
                                            group_name = "Cephalopoda",
                                            strata_name = "ntr")

res_cephall_ntr <- determine_xmin_and_fit_mlebins(data_cephall_ntr)

res_cephall_ntr$mlebins_fit
```



```{r, cephallntrplot}
plot(res_cephall_ntr,
     xlim_hist = c(0, 60),
     main = "Cephalopoda (all) NTR")
```

Now to compare the three strata for Cephalopoda (all).

```{r, cephallsumm}
res_table_cephall <- dplyr::tibble()
res_table_cephall <- add_row_to_summary_table(res_table_cephall,
                                      res_cephall_baseline,
                                      "Cephalopoda all",
                                      "Baseline")

res_table_cephall <- add_row_to_summary_table(res_table_cephall,
                                      res_cephall_fg,
                                      "Cephalopoda all",
                                      "fg")
res_table_cephall <- add_row_to_summary_table(res_table_cephall,
                                      res_cephall_ntr,
                                      "Cephalopoda all",
                                      "ntr")
```

Results combined are:

```{r, rescephalltable}
knitr::kable(res_table_cephall, digits = 2)
```



## Global xmin and xmax for Cephalopoda (all)

So doing the modal approach gave different xmin between baseline and other two;
should go with the largest of those. Though might give quite shallow power law
for baseline.

Alternative is to use xmin for the aggregated data, TODO try that also.

xmax was set independently, so we should select the globally largest for
this species group.

```{r, cephallxmax}
cephall_xmin <- max(res_table_cephall$x_min)
cephall_xmin
cephall_xmax <- max(res_table_cephall$x_max)
cephall_xmax
```

## Cephalopoda (all) baseline with fixed xmin and xmax

```{r, cephallbaselinexmax}
res_cephall_baseline_fix_xmax <-
  determine_xmin_and_fit_mlebins(data_cephall_baseline,
                                 x_min = cephall_xmin,
                                 x_max = cephall_xmax)
```



```{r, cephallbaselineplotxmax}
plot(res_cephall_baseline_fix_xmax,
     xlim_hist = c(0, 60),
     main = "Cephalopoda (all) baseline")
```

## Cephalopoda (all) FG with fixed xmin and xmax

```{r, cephallfgxmax}
res_cephall_fg_fix_xmax <- determine_xmin_and_fit_mlebins(data_cephall_fg,
                                                        x_min = cephall_xmin,
                                                        x_max = cephall_xmax)
```



```{r, cephallfgplotxmax}
plot(res_cephall_fg_fix_xmax,
     xlim_hist = c(0, 60),
     main = "Cephalopoda (all) FG")
```

## Cephalopoda (all) NTR with fixed xmin and xmax

```{r, cephallbaelinexmax}
res_cephall_ntr_fix_xmax <- determine_xmin_and_fit_mlebins(data_cephall_ntr,
                                                         x_min = cephall_xmin,
                                                         x_max = cephall_xmax)
```

```{r, cephallntrplotxmax}
plot(res_cephall_ntr_fix_xmax,
     xlim_hist = c(0, 60),
     main = "Cephalopoda (all) NTR")
```

Now to compare the three strata for Cephalopoda (all) with xmax fixed:

```{r, cephallsummfix}
res_table_cephall_fix_xmax <- dplyr::tibble()
res_table_cephall_fix_xmax <- add_row_to_summary_table(res_table_cephall_fix_xmax,
                                      res_cephall_baseline_fix_xmax,
                                      "Cephalopoda all",
                                      "Baseline")

res_table_cephall_fix_xmax <- add_row_to_summary_table(res_table_cephall_fix_xmax,
                                      res_cephall_fg_fix_xmax,
                                      "Cephalopoda all",
                                      "fg")
res_table_cephall_fix_xmax <- add_row_to_summary_table(res_table_cephall_fix_xmax,
                                      res_cephall_ntr_fix_xmax,
                                      "Cephalopoda all",
                                      "ntr")
```

Results combined are:

```{r, rescephalltablexmax}
knitr::kable(res_table_cephall_fix_xmax, digits = 2)
```

## Determine cephalopod small/large cutoff

Now repeat the above for small and large separately, since the
first baseline plot clearly suggests a natural break. Explicitly this is at
20.89, with fg and ntr saying maybe something similar. 24 might be a sensible
consistent way of breaking up the data, but 20.89 seems more
obvious and occurs for all (so makes sense to use it for xmin).
Specifically 20.892555.... Here are all the data:

```{r, cephallbaselinebreak}
dplyr::arrange(data_cephall_baseline, bin_min) %>% a()

dplyr::arrange(data_cephall_fg, bin_min) %>% a()

dplyr::arrange(data_cephall_ntr, bin_min) %>% a()

# And the rows in each for that xmin suggestion are:
ceph_small_large_delineation <- dplyr::filter(data_cephall_baseline,
                                              species == "Illex coindetii",
                                              bin_min > 20,
                                              bin_min < 21)$bin_min  # to get
                                              # the exact value

ceph_small_large_delineation

dplyr::filter(data_cephall_baseline,
              species == "Illex coindetii",
              bin_min == ceph_small_large_delineation)
dplyr::filter(data_cephall_fg,
              species == "Illex coindetii",
              bin_min == ceph_small_large_delineation)
dplyr::filter(data_cephall_ntr,
              species == "Illex coindetii",
              bin_min == ceph_small_large_delineation)
```

Okay, use that for delineating small/big organisms. Make the xmax for small ones
a little less than that to ensure the cutoff is correct.


## Global xmin and xmax for Cephalopoda (small)


```{r, cephsmallxmax}
cephsmall_xmin <- max(res_table_cephall$x_min)
cephsmall_xmin
cephsmall_xmax <- floor(ceph_small_large_delineation*100)/100   # rounding down to 0.01
cephsmall_xmax
```

## Cephalopoda (small) baseline with fixed xmin and xmax

```{r, cephsmallbaselinexmax}
data_cephsmall_baseline <- dplyr::filter(data_cephall_baseline,
                                         bin_min < cephsmall_xmax)
res_cephsmall_baseline_fix_xmax <-
  determine_xmin_and_fit_mlebins(data_cephsmall_baseline,
                                 x_min = cephsmall_xmin,
                                 x_max = cephsmall_xmax)
```



```{r, cephsmallbaselineplotxmax}
plot(res_cephsmall_baseline_fix_xmax,
     xlim_hist = c(0, 21),
     main = "Cephalopoda (small) baseline")
```

## Cephalopoda (small) FG with fixed xmin and xmax

```{r, cephsmallfgxmax}
data_cephsmall_fg <- dplyr::filter(data_cephall_fg,
                                   bin_min < cephsmall_xmax)

res_cephsmall_fg_fix_xmax <- determine_xmin_and_fit_mlebins(data_cephsmall_fg,
                                                            x_min = cephsmall_xmin,
                                                            x_max = cephsmall_xmax)
```



```{r, cephsmallfgplotxmax}
plot(res_cephsmall_fg_fix_xmax,
     xlim_hist = c(0, 21),
     main = "Cephalopoda (small) FG")
```

## Cephalopoda (small) NTR with fixed xmin and xmax

```{r, cephsmallbaelinexmax}
data_cephsmall_ntr <- dplyr::filter(data_cephall_ntr,
                                    bin_min < cephsmall_xmax)

res_cephsmall_ntr_fix_xmax <- determine_xmin_and_fit_mlebins(data_cephsmall_ntr,
                                                             x_min = cephsmall_xmin,
                                                             x_max = cephsmall_xmax)
```

```{r, cephsmallntrplotxmax}
plot(res_cephsmall_ntr_fix_xmax,
     xlim_hist = c(0, 21),
     main = "Cephalopoda (small) NTR")
```

Now to compare the three strata for Cephalopoda (small) with common xmin and
xmax:

```{r, cephsmallsummfix}
res_table_cephsmall_fix_xmax <- dplyr::tibble()
res_table_cephsmall_fix_xmax <- add_row_to_summary_table(res_table_cephsmall_fix_xmax,
                                      res_cephsmall_baseline_fix_xmax,
                                      "Cephalopoda small",
                                      "Baseline")

res_table_cephsmall_fix_xmax <- add_row_to_summary_table(res_table_cephsmall_fix_xmax,
                                      res_cephsmall_fg_fix_xmax,
                                      "Cephalopoda small",
                                      "fg")
res_table_cephsmall_fix_xmax <- add_row_to_summary_table(res_table_cephsmall_fix_xmax,
                                      res_cephsmall_ntr_fix_xmax,
                                      "Cephalopoda small",
                                      "ntr")
```

Results combined are:

```{r, rescephsmalltablexmax}
knitr::kable(res_table_cephsmall_fix_xmax, digits = 2)
```

## Global xmin and xmax for Cephalopoda (large)

```{r, cephlargexmax}
cephlarge_xmin <- ceph_small_large_delineation
cephlarge_xmin
cephlarge_xmax <- cephall_xmax
cephlarge_xmax
```

## Cephalopoda (large) baseline with fixed xmin and xmax

```{r, cephlargebaselinexmax}
data_cephlarge_baseline <- dplyr::filter(data_cephall_baseline,
                                         bin_min > cephsmall_xmax)
res_cephlarge_baseline_fix_xmax <-
  determine_xmin_and_fit_mlebins(data_cephlarge_baseline,
                                 x_min = cephlarge_xmin,
                                 x_max = cephlarge_xmax)
```



```{r, cephlargebaselineplotxmax}
plot(res_cephlarge_baseline_fix_xmax,
     xlim_hist = c(0, 100),
     main = "Cephalopoda (large) baseline")
```

## Cephalopoda (large) FG with fixed xmin and xmax

```{r, cephlargefgxmax}
data_cephlarge_fg <- dplyr::filter(data_cephall_fg,
                                   bin_min > cephsmall_xmax)

res_cephlarge_fg_fix_xmax <- determine_xmin_and_fit_mlebins(data_cephlarge_fg,
                                                            x_min = cephlarge_xmin,
                                                            x_max = cephlarge_xmax)
```



```{r, cephlargefgplotxmax}
plot(res_cephlarge_fg_fix_xmax,
     xlim_hist = c(0, 100),
     main = "Cephalopoda (large) FG")
```

## Cephalopoda (large) NTR with fixed xmin and xmax

```{r, cephlargebaelinexmax}
data_cephlarge_ntr <- dplyr::filter(data_cephall_ntr,
                                    bin_min > cephsmall_xmax)

res_cephlarge_ntr_fix_xmax <- determine_xmin_and_fit_mlebins(data_cephlarge_ntr,
                                                             x_min = cephlarge_xmin,
                                                             x_max = cephlarge_xmax)
```

```{r, cephlargentrplotxmax}
plot(res_cephlarge_ntr_fix_xmax,
     xlim_hist = c(0, 100),
     main = "Cephalopoda (large) NTR")
```

Now to compare the three strata for Cephalopoda (large) with common xmin and
xmax:

```{r, cephlargesummfix}
res_table_cephlarge_fix_xmax <- dplyr::tibble()
res_table_cephlarge_fix_xmax <- add_row_to_summary_table(res_table_cephlarge_fix_xmax,
                                      res_cephlarge_baseline_fix_xmax,
                                      "Cephalopoda large",
                                      "Baseline")

res_table_cephlarge_fix_xmax <- add_row_to_summary_table(res_table_cephlarge_fix_xmax,
                                      res_cephlarge_fg_fix_xmax,
                                      "Cephalopoda large",
                                      "fg")
res_table_cephlarge_fix_xmax <- add_row_to_summary_table(res_table_cephlarge_fix_xmax,
                                      res_cephlarge_ntr_fix_xmax,
                                      "Cephalopoda large",
                                      "ntr")
```

Results combined are:

```{r, rescephlargetablexmax}
knitr::kable(res_table_cephlarge_fix_xmax, digits = 2)
```

# Combining results


HERE written this but not tested yet, since main code is still running. Maybe
should save the required results here and make a new file to do the aggregation
and the main plots.

For each of baseline, FG, and NTR, can we produce an aggregated plot. Based on
what I figured out in 'Aggregating size spectra' document. Here just focus on
developing the code, based somewhat on that and code developed then.

First, let's create a table of the main results from above that we want to move
forward with. Just from the `res_table_***` objects. Then we'll need the raw
data from each in turn. Could then functionalise this maybe, but probably best
to keep explicit and tailored for different applications.

```{r, aggregatetable}
res_table_agg <- rbind(res_table_crust_fix_xmax,
                       res_table_actin_fix_xmax,
                       res_table_chondr_fix_xmax2,
                       res_table_cephsmall_fix_xmax,
                       res_table_cephlarge_fix_xmax)


knitr::kable(res_table_agg,
             digits = 2)
```


## Aggregated fit to data

```{r, aggregatefit, eval = FALSE}
baseline_agg <- aggregate_mlebins(list(crust = res_crust_baseline_fix_xmax,
                                       actin = res_actin_baseline_fix_xmax,
                                       chondr = res_chondr_baseline_fix_xmax2,
                                       cephsmall = res_cephsmall_baseline_fix_xmax,
                                       cephlarge = res_cephlarge_baseline_fix_xmax),
                                  res_table_agg,
                                  strata = "Baseline")

fg_agg <- aggregate_mlebins(list(crust = res_crust_fg_fix_xmax,
                                 actin = res_actin_fg_fix_xmax,
                                 chondr = res_chondr_fg_fix_xmax2,
                                 cephsmall = res_cephsmall_fg_fix_xmax,
                                 cephlarge = res_cephlarge_fg_fix_xmax),
                            res_table_agg,
                            strata = "fg")

ntr_agg <- aggregate_mlebins(list(crust = res_crust_ntr_fix_xmax,
                                  actin = res_actin_ntr_fix_xmax,
                                  chondr = res_chondr_ntr_fix_xmax2,
                                  cephsmall = res_cephsmall_ntr_fix_xmax,
                                  cephlarge = res_cephlarge_ntr_fix_xmax),
                             res_table_agg,
                             strata = "ntr")
```

Each one is saved as a `mlebins_list object` to then create plotting function.



## Saving what is needed for plots and tables, and doing in zabala-summary.Rmd

```{r, saving}
saveRDS(baseline_agg,
        "baseline_agg.rds")
saveRDS(fg_agg,
        "fg_agg.rds")
saveRDS(ntr_agg,
        "ntr_agg.rds")
saveRDS(res_table_agg,
        "res_table_agg.rds")
```

The current file is very big and takes a while to run (also .pdf keeps hanging),
so time to save what we need and do a clean one.
